---
title: "Live demo 1"
author: "Andrew Irwin"
date: "2022-09-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

* Overview of main components of Rstudio
  * Editing window
  * Console
  * Terminal
  * Files
  * Plots 
  * Help
  * Environment
  * Packages
* Overview of Markdown
  * Headings
  * Code blocks
  * Knitting
  * Quarto
* R projects
  * What's a file?
  * How to name files
* Making and revising notes. Learning and not getting lost.

### Installing packages

```{r eval=FALSE}
my_packages <- c("tidyverse", "broom", "coefplot", "cowplot",
                 "gapminder", "GGally", "ggrepel", "ggridges", "gridExtra",
                 "here", "interplot", "margins", "maps", "mapproj",
                 "mapdata", "MASS", "quantreg", "rlang", "scales",
                 "survey", "srvyr", "viridis", "viridisLite", 
                 "socviz", "devtools", "patchwork", "usethis", "gitcreds",
                 "remotes", "paletteer", "ggthemes", "ggtext")

install.packages(my_packages)
```

## Cloud services

* rstudio.cloud

## Why these tools and not some others?

* Python
* Excel, Office suite

Why is programming a computer worth the effort to learn, to use?

What about learning fundamentals instead of the specifics of one particular software package?


## Other topics get their own lessons

* Grammar of Graphics & ggplot2
* Tidy data: dplyr, tidyverse packages
* Version control: git and github
