---
title: "Demo"
author: "Andrew Irwin"
date: 2026-01-13
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

-   Overview of main components of Rstudio
    -   Editing window
    -   Console
    -   Terminal
    -   Files
    -   Plots
    -   Help
    -   Environment
    -   Packages
-   Overview of Markdown
    -   Headings
    -   Code blocks
    -   Knitting
    -   Quarto
-   R projects
    -   What's a file?
    -   How to name files
-   Making and revising notes. Learning and not getting lost.

### Installing packages

```{r eval=FALSE}
my_packages <- c("tidyverse", "broom", "coefplot", "cowplot",
                 "gapminder", "GGally", "ggrepel", "ggridges", "gridExtra",
                 "here", "interplot", "margins", "maps", "mapproj",
                 "mapdata", "MASS", "quantreg", "rlang", "scales",
                 "survey", "srvyr", "viridis", "viridisLite", 
                 "socviz", "devtools", "patchwork", "usethis", "gitcreds",
                 "remotes", "paletteer", "ggthemes", "ggtext")

install.packages(my_packages)
```

(Sometimes I set `eval=FALSE` on a chunk so that the code is displayed but not executed. Take note and change as needed.)

## Cloud services

-   [posit.cloud](https://posit.cloud/)

## Why these tools and not some others?

-   Python
-   JavaScript
-   Excel, Office suite

Why is programming a computer worth the effort to learn, to use?

What about learning fundamentals instead of the specifics of one particular software package?

## Other topics get their own lessons

-   Grammar of Graphics & ggplot2
-   Tidy data: dplyr, tidyverse packages
-   Version control: git and github
