---
title: "First Plot demonstration"
author: "Andrew Irwin"
date: "2022-09-13"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(gapminder)
```

This is a demonstration showing how to make your first plot using R, Rstudio, and ggplot.

Steps:

* Be sure you are using the `tidyverse` and `gapminder` libraries (and that they are installed on your computer).
* Create a ggplot using the data `gapminder` and a mapping between data and aesthetic features. For example put GDP per capita on the x axis and life expectancy on the y axis.
* Make a scatterplot using `geom_point`

```{r}

```


```{r}

```




Refine your plot:

* Add axis labels using `labs` and the arguments `x`, `y`, `title`, `subtitle`, `caption`, `tag` (some of these are optional and not always needed.)
* Use `theme` to make the text bigger by setting the `size` property of `text` with the `element_text` function.
* Add colour to the plot. Change the label for your colour scale.
* Transform the x axis to be on a log scale using `scale_x_log10`

```{r}



```

```{r}


```

```{r}


```


Repeat:

* Do this exercise with another dataset such as `mtcars` or `palmerpenguins`.

Refer to the slides for this lesson for more ideas and examples.
