---
title: "First Plot demonstration"
author: "Andrew Irwin"
date: "2022-09-13"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(gapminder)
```

This is a demonstration showing how to make your first plot using R, Rstudio, and ggplot.

Steps:

* Be sure you are using the `tidyverse` and `gapminder` libraries (and that they are installed on your computer).
* Create a ggplot using the data `gapminder` and a mapping between data and aesthetic features. For example put GDP per capita on the x axis and life expectancy on the y axis.
* Make a scatterplot using `geom_point`

```{r}
ggplot(gapminder, aes(x = gdpPercap, y = lifeExp)) + geom_point()
```


```{r}
ggplot(gapminder, aes(y = gdpPercap, x = lifeExp)) + geom_point()
```




Refine your plot:

* Add axis labels using `labs` and the arguments `x`, `y`, `title`, `subtitle`, `caption`, `tag` (some of these are optional and not always needed.)
* Use `theme` to make the text bigger by setting the `size` property of `text` with the `element_text` function.
* Add colour to the plot. Change the label for your colour scale.
* Transform the x axis to be on a log scale using `scale_x_log10`

```{r}
ggplot(gapminder, aes(y = gdpPercap, 
                      x = lifeExp)) + 
  geom_point() +
  labs(x = "GDP per capita (US$, inflation adjusted",
       y = "Life expectancy (years, at birth)",
       title = "Life expectancy varies with national income",
       subtitle = "Something more here",
       caption = "Source: gapminder")

```

```{r}
ggplot(gapminder, aes(y = gdpPercap, 
                      x = lifeExp,
                      color = continent),
       ) + 
  geom_point() +
  labs(y = "GDP per capita (US$, inflation adjusted",
       x = "Life expectancy (years, at birth)",
       title = "Life expectancy varies with national income",
       subtitle = "Something more here",
       caption = "Source: gapminder")

```

```{r}
ggplot(gapminder, aes(y = gdpPercap, 
                      x = lifeExp,
                      color = continent),
       ) + 
  geom_point() +
  labs(y = "GDP per capita (US$, inflation adjusted",
       x = "Life expectancy (years, at birth)",
       title = "Life expectancy varies with national income",
       subtitle = "Something more here",
       caption = "Source: gapminder") +
  scale_y_log10() +
  theme(text = element_text(size=14))

```


Repeat:

* Do this exercise with another dataset such as `mtcars` or `palmerpenguins`.

Refer to the slides for this lesson for more ideas and examples.
