---
title: "Grammar of Graphics practice"
author: "Andrew Irwin"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    toc-location: left       # 'right' (default) or 'left'
    toc-float: true          # Ensures the TOC moves as you scroll
    embed-resources: true    # Embeds images/css for a single file
---

```{r}
library(tidyverse)
library(socviz) # data
library(datasets) # data
library(palmerpenguins) # data
library(gapminder) # data
```

Practice making plots of these data: `gss_sm`, `gss_lon`, `penguins`, `penguins_raw`, `gapminder`, and many others in the `datasets` package.

Examine the data using `View`, `glimpse`. Read the help pages about each dataset.

```{r}
glimpse(ToothGrowth)
glimpse(gss_sm)
glimpse(penguins)
```

## Aesthetic mappings to use

x, y, color, fill, shape, size, group

## Geometries to try

* geom_bar, geom_count,
* geom_point, geom_line, geom_boxplot, geom_jitter, 
* geom_histogram, geom_density,
* geom_hline, geom_vline,  geom_rug,
* stat_summary,
* geom_function

Which should you use with categorical data? quantitative data? A mixture?

What happens if you use the wrong kind of data?

Take a look at the help for each geom using the Help panel in Rstudio or on the ggplot
[website](https://ggplot2.tidyverse.org/index.html).

Try a few simple themes:

* theme_bw(base_size = 14)
* theme_minimal(base_family = "serif") # sans, mono, font name
* theme_classic()

As we practice, we will describe each geom here and provide some examples.

## Examples

Created in class using the suggestions above.

```{r warning=FALSE, message=FALSE}
#| fig.height: 4.25
gss_sm |> ggplot(aes(y = region)) + geom_bar() + 
  theme_bw(base_size=14)
```

```{r warning=FALSE, message=FALSE}
#| fig.height: 4.25
#| fig.align: "center"
gss_sm |> ggplot() +
  geom_bar(aes(y = region, fill = degree ))
gss_sm |> ggplot() +
  geom_bar(aes(y = region, colour = degree ))
gss_sm |> ggplot() +
  geom_bar(aes(y = region, fill = degree ), color = "white")
gss_sm |> ggplot() +
  geom_bar(aes(y = region, fill = degree, color = "white"))
gss_sm |> ggplot() +
  geom_bar(aes(y = degree, fill = region ))

```


```{r warning=FALSE, message=TRUE}
#| fig.height: 4
#| fig.align: "center"
#| dpi: 150
penguins |> ggplot() + 
  geom_histogram(aes(body_mass_g)) 
penguins |> ggplot() + 
  geom_histogram(aes(body_mass_g), binwidth = 225) 
penguins |> ggplot() + 
  geom_histogram(aes(y = body_mass_g)) 
penguins |> ggplot() + 
  geom_histogram(aes(y = body_mass_g, fill = species)) 
penguins |> ggplot() + 
  geom_histogram(aes(x = body_mass_g, fill = species), color = "white") 
```


```{r warning=FALSE, message=FALSE}
#| fig.height: 3
#| fig.align: "center"
#| dpi: 150
penguins |> ggplot() + 
  geom_density(aes(flipper_length_mm)) 
penguins |> ggplot() + 
  geom_density(aes(flipper_length_mm), bw = 2) 
penguins |> ggplot() + 
  geom_density(aes(x = body_mass_g, fill = species)) 
penguins |> ggplot() + 
  geom_density(aes(x = body_mass_g, fill = species), alpha = 0.4) 
penguins |> ggplot() + 
  geom_density(aes(x = body_mass_g, colour = species)) 
```



```{r warning=TRUE, message=FALSE}
#| fig.height: 3
#| fig.align: "center"
#| dpi: 150
penguins |> ggplot() + 
  geom_boxplot(aes(body_mass_g)) 
penguins |> ggplot() + 
  geom_boxplot(aes(y = body_mass_g)) 
penguins |> ggplot() + 
  geom_boxplot(aes(y = body_mass_g, x = species, fill = species)) 
penguins |> ggplot() + 
  geom_boxplot(aes(y = body_mass_g, x = species, fill = species), color = "white") 
```

```{r}
penguins |> ggplot() + 
  geom_count(aes(species, island))
```

```{r}
penguins |> ggplot() + 
  geom_bar(aes(species, fill = sex))
```

```{r}
gapminder |> ggplot() + 
  geom_point(aes(x = gdpPercap, y = lifeExp, size = pop, color = continent)) +
  theme_minimal()
gapminder |> ggplot() + 
  geom_point(aes(x = log10(gdpPercap), y = lifeExp, size = pop, color = continent)) +
  theme_minimal()
gapminder |> ggplot() + 
  geom_point(aes(x = gdpPercap, y = lifeExp, size = pop, color = continent)) +
  scale_x_log10() + 
  theme_minimal()
gapminder |> filter(year == 2002) |>
  ggplot() + 
  geom_point(aes(x = gdpPercap, y = lifeExp, size = log10(pop), color = continent)) +
  scale_x_log10() + 
  theme_minimal()
```


```{r}
gapminder |> filter(country == "Canada") |>
  ggplot() + 
  geom_line(aes(x = year, y = lifeExp, color = continent)) +
  theme_minimal()
gapminder |> filter(country == "Canada") |>
  ggplot() + 
  geom_line(aes(x = year, y = lifeExp, color = country)) +
  ylim(0, NA) + 
  theme_minimal()
gapminder |> filter(country == "Canada"  | country == "China") |>
  ggplot() + 
  geom_line(aes(x = year, y = lifeExp, color = country)) +
  ylim(0, NA) + 
  theme_minimal()
gapminder |> filter(country %in% c("Canada", "China", "Brazil", "United States")) |>
  ggplot() + 
  geom_line(aes(x = year, y = lifeExp, color = country), linewidth = 1.2) +
  geom_point(aes(x = year, y = lifeExp, color = country), size = 4, shape = "o") +
  ylim(0, NA) + 
  theme_minimal() +
  labs(x = "Year", y = "Life Expectancy at birth", 
       colour = "Country",
       title = "Life expectancy over time",
       subtitle = "From gapminder")
```

## Previous year's explorations

### An example to warm up with

```{r warning=FALSE, message=FALSE, fig.height=4.25, fig.align="center", dpi=150}
gss_sm |> ggplot(aes(y = region, color = sex)) + geom_bar()
gss_sm |> ggplot(aes(y = region, fill = sex)) + geom_bar()
```


```{r}
glimpse(penguins)
```

```{r}
penguins |> ggplot(aes(x = flipper_length_mm)) + geom_histogram(binwidth = 5)
```

```{r}
penguins |> ggplot(aes(x = flipper_length_mm, fill = species)) + geom_histogram(binwidth = 5)
penguins |> ggplot(aes(x = flipper_length_mm, fill = island)) + geom_histogram(binwidth = 5)
```


### Make lots of your own experiments

Can we show two categorical variables on one plot?

```{r}
penguins |> ggplot(aes(species, island)) + geom_count()
penguins |> ggplot(aes(species, island)) + geom_jitter()
penguins |> ggplot(aes(species, island)) + geom_jitter(aes(color = sex)) # a bit random for my taste
penguins |> ggplot(aes(species, island)) + geom_count(aes(color = sex)) # not sure I understand this one

```

### Make the plots look better

* Add labels with `labs(x = "...", y = "...", title = "...", color = "...", fill = "...")`.

```{r}
penguins |> ggplot(aes(species, island, color = sex)) + geom_jitter() +
  labs(x = "Species", y = "Island", color = "Sex")

```

* Make the text bigger with `theme(text = element_text(size = 15))`.

* Turn off the grey shading with `theme_bw()`

* Always use `+` to "add" parts of a ggplot together.

### gapminder experiments

```{r}
str(gapminder)
```

```{r}
gapminder |> ggplot(aes(x = gdpPercap, y = lifeExp )) + geom_point() +
  labs(x = "GDP per capita ($)", y = "Life expectancy (years)")
```

```{r}
gapminder |> ggplot(aes(x = gdpPercap, y = lifeExp,
                        color = year, shape = continent)) + geom_point() +
  labs(x = "GDP per capita ($)", y = "Life expectancy (years)")
gapminder |> ggplot(aes(x = log10(gdpPercap), y = lifeExp,
                        color = year, shape = continent)) + geom_point() +
  labs(x = "GDP per capita ($)", y = "Life expectancy (years)")
```

Can we see how GDP per capita varies with year? And maybe continent too?

```{r}
gapminder |> ggplot(aes(y = log10(gdpPercap), x = year)) + geom_point()
gapminder |> ggplot(aes(y = log10(gdpPercap), x = year)) + geom_boxplot() # ugh!!
gapminder |> ggplot(aes(y = log10(gdpPercap), x = factor(year))) + geom_boxplot()
gapminder |> ggplot(aes(y = log10(gdpPercap), x = factor(year), fill = continent)) + geom_boxplot()
gapminder |> ggplot(aes(y = log10(gdpPercap), fill = factor(year), x = continent)) + geom_boxplot()
```


### Examples from previous courses

Left here as more examples.

```{r}
penguins |> ggplot(aes(x = sex)) + geom_bar()
penguins |> ggplot(aes(x = sex, color = species)) + geom_bar()
penguins |> ggplot(aes(x = sex, fill = species)) + geom_bar()
penguins |> ggplot(aes(fill = sex, x = species)) + geom_bar()
penguins |> ggplot(aes(fill = sex, x = species)) + geom_bar(position = "dodge") 
penguins |> ggplot(aes(x = species, y = sex)) + geom_count()
# penguins |> ggplot(aes(x = species)) + geom_count() # fails can't drqw with only one of x or y
penguins |> ggplot(aes(x = species, y = island, color = sex)) + geom_count(position = "jitter") # terrible, but fun!
penguins |> ggplot(aes(flipper_length_mm, bill_length_mm)) + geom_point()
penguins |> ggplot(aes(flipper_length_mm, bill_length_mm, color = species)) + geom_point()
penguins |> ggplot(aes(flipper_length_mm, bill_length_mm, color = sex)) + geom_point()
penguins |> na.omit() |> ggplot(aes(flipper_length_mm, bill_length_mm, color = species, size = sex)) + geom_point()
penguins |> ggplot(aes(body_mass_g, bill_length_mm)) + geom_line() # works, but silly
gapminder |> filter(country %in% c("China", "Canada", "India", "France", "Argentina", "Lybia")) |>
  ggplot(aes(year, lifeExp)) + geom_line()
gapminder |> filter(country %in% c("China", "Canada", "India", "France", "Argentina", "Lybia")) |>
  ggplot(aes(year, lifeExp, color = country)) + geom_line()
gapminder |> 
  ggplot(aes(year, lifeExp, group = country)) + geom_line()
gapminder |> ggplot(aes(x = continent, y = lifeExp)) + geom_boxplot()
gapminder |> ggplot(aes(x = year, y = lifeExp)) + geom_boxplot()  # doesn't work the way I hoped!
gapminder |> ggplot(aes(x = factor(year), y = lifeExp)) + geom_boxplot()
gapminder |> ggplot(aes(x = factor(year), y = lifeExp, color = continent)) + geom_boxplot()
gapminder |> ggplot(aes(lifeExp, fill = continent)) + geom_histogram()
gapminder |> ggplot(aes(lifeExp, color = continent)) + geom_density()
gapminder |> ggplot(aes(y = lifeExp, x = continent, color = factor(year))) + stat_summary() +
  labs(x = "Continent", y = "Life Expectancy (years)", color = "Year") +
  theme_bw()
```
