---
title: "Grammar of Graphics practice"
author: "Andrew Irwin"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(socviz) # data
library(datasets) # data
library(palmerpenguins) # data
library(gapminder)
```

Practice making plots of these data: `gss_sm`, `gss_lon`, `penguins`, 
`penguins_raw`, `gapminder`, and many others in the `datasets` package.

Examine the data using `View`, `glimpse`.

```{r}
glimpse(ToothGrowth)
```

## Aesthetic mappings to use

x, y, color, fill, shape, size, group

## Geometries to try

* geom_bar, geom_count,
* geom_point, geom_line, geom_boxplot, geom_jitter, 
* geom_histogram, geom_density,
* geom_hline, geom_vline,  geom_rug,
* stat_summary,
* geom_function

Which should you use with categorical data? quantitative data? A mixture?
What happens if you use the wrong kind of data?

Take a look at the help for each geom using the Help panel in Rstudio or on the ggplot
[website](https://ggplot2.tidyverse.org/index.html).

As we practice, we will describe each geom here and provide some examples.

## An example to warm up with

```{r}
penguins |> ggplot(aes(x = flipper_length_mm)) + geom_histogram()
```

## Make lots of your own experiments

```{r}


```

## Make the plots look better

* Add labels with `labs(x = "...", y = "...", title = "...", color = "...", fill = "...")`.

* Make the text bigger with `theme(text = element_text(size = 15))`.

* Turn off the grey shading with `theme_bw()`

* Always use `+` to "add" parts of a ggplot together.


## Examples from previous courses

Left here as more examples.

```{r}
penguins |> ggplot(aes(x = sex)) + geom_bar()
penguins |> ggplot(aes(x = sex, color = species)) + geom_bar()
penguins |> ggplot(aes(x = sex, fill = species)) + geom_bar()
penguins |> ggplot(aes(fill = sex, x = species)) + geom_bar()
penguins |> ggplot(aes(fill = sex, x = species)) + geom_bar(position = "dodge") 
penguins |> ggplot(aes(x = species, y = sex)) + geom_count()
# penguins |> ggplot(aes(x = species)) + geom_count() # fails can't drqw with only one of x or y
penguins |> ggplot(aes(x = species, y = island, color = sex)) + geom_count(position = "jitter") # terrible, but fun!
penguins |> ggplot(aes(flipper_length_mm, bill_length_mm)) + geom_point()
penguins |> ggplot(aes(flipper_length_mm, bill_length_mm, color = species)) + geom_point()
penguins |> ggplot(aes(flipper_length_mm, bill_length_mm, color = sex)) + geom_point()
penguins |> na.omit() |> ggplot(aes(flipper_length_mm, bill_length_mm, color = species, size = sex)) + geom_point()
penguins |> ggplot(aes(body_mass_g, bill_length_mm)) + geom_line() # works, but silly
gapminder |> filter(country %in% c("China", "Canada", "India", "France", "Argentina", "Lybia")) |>
  ggplot(aes(year, lifeExp)) + geom_line()
gapminder |> filter(country %in% c("China", "Canada", "India", "France", "Argentina", "Lybia")) |>
  ggplot(aes(year, lifeExp, color = country)) + geom_line()
gapminder |> 
  ggplot(aes(year, lifeExp, group = country)) + geom_line()
gapminder |> ggplot(aes(x = continent, y = lifeExp)) + geom_boxplot()
gapminder |> ggplot(aes(x = year, y = lifeExp)) + geom_boxplot()  # doesn't work the way I hoped!
gapminder |> ggplot(aes(x = factor(year), y = lifeExp)) + geom_boxplot()
gapminder |> ggplot(aes(x = factor(year), y = lifeExp, color = continent)) + geom_boxplot()
gapminder |> ggplot(aes(lifeExp, fill = continent)) + geom_histogram()
gapminder |> ggplot(aes(lifeExp, color = continent)) + geom_density()
gapminder |> ggplot(aes(y = lifeExp, x = continent, color = factor(year))) + stat_summary() +
  labs(x = "Continent", y = "Life Expectancy (years)", color = "Year") +
  theme_bw()
```
