---
title: "Grammar of Graphics practice"
author: "Andrew Irwin"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(socviz) # data
library(datasets) # data
library(palmerpenguins) # data
```

Practice making plots of these data: `gss_sm`, `gss_lon`, `penguins`, 
`penguins_raw`, `gapminder`, and many others in the `datasets` package.

Examine the data using `View`, `glimpse`.

## Aesthetic mappings to use

x, y, color, fill, shape, size, group

## Geometries to try

geom_point, geom_line, geom_boxplot, geom_jitter, geom_histogram, geom_density,
geom_function, geom_bar, geom_hline, geom_vline, geom_count, geom_rug,
stat_summary

Which should you use with categorical data? quantitative data? A mixture? What happens
if you use the wrong kind of data?

Take a look at the help for each geom using the Help panel in Rstudio or on the ggplot
[website](https://ggplot2.tidyverse.org/index.html).

As we practice, we will describe each geom here and provide some examples.

## An example to warm up with

```{r}
penguins |> ggplot(aes(y = flipper_length_mm)) + geom_histogram()
```

## Make the plots look better

Add labels with `labs(x = "...", y = "...", title = "...", color = "...", fill = "...")`.

Make the text bigger with `theme(text = element_text(size = 15))`.

Turn of the grey shading with `theme_bw()`

Always use `+` to "add" parts of a ggplot together.

