---
title: "Summarizing data practice"
author: "AJ Irwin"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
```


Use the dplyr functions: mutate, filter, group_by, summarize, select, arrange to answer the following questions.

How many kinds of species of penguin are in the `penguins`? How many observations of each species?

```{r}

```

What is the maximum and minimum flipper length?

```{r}

```

Find the observation with the largest flipper length.

```{r}

```

Make a table showing the number of observations of each combination of species and island.

```{r}

```

Compute the ratio of the bill length to bill depth and make a histogram with these data.

```{r}

```

Repeat that calculation for each species.

```{r}

```

Compute the mean and stanard deviation of body mass for each species and sex.

```{r}

```

Decode this complex calculation from the course notes by executing just one line at a time, making a prediction about the result, and checking your understanding. Change the calculation to make it do something similar but different to test your understanding. (For example, try using `slice_tail` instead of `slice_head`, or remove the `-` in the `arrange` function.)

```{r}
diamonds |>
  mutate(price_per_carat = price / carat) |>
  group_by(color, clarity, cut) |>
  summarise(median_price_per_carat = median(price_per_carat), 
            n = n(),
            .groups = "drop") |>
  arrange(-median_price_per_carat) |>
  group_by(cut) |>
  slice_head(n=2) |>
  arrange(-median_price_per_carat)
```

