---
title: "Summarizing data practice"
author: "AJ Irwin"
date: "2024-02-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
```

From slides

```{r}
ChickWeight |> 
  group_by(Chick) |>
  mutate(growth = (weight-min(weight)) / Time)
```

Try single = for logical comparison and see what happens!

```{r error=TRUE}
ChickWeight |> 
  filter(Time == 10) |> 
  group_by(Diet) |> 
  summarize(weight_10 = mean(weight), .groups = "drop")
```


Use the dplyr functions: mutate, filter, group_by, summarize, select, arrange to answer the following questions.

* How many kinds of species of penguin are in the dataset `penguins`? How many observations of each species?

```{r}
penguins |> count(species)
```


* What is the maximum and minimum flipper length?

```{r}
penguins |> summarize(max_flipper = max(flipper_length_mm),
                      min_flipper = min(flipper_length_mm))
penguins |> summarize(max_flipper = max(flipper_length_mm, na.rm=TRUE),
                      min_flipper = min(flipper_length_mm, na.rm=TRUE),
                      count = n())

```

* Use `na.omit` and `filter(!is.na())` to remove missing data.

```{r}
penguins |> filter(!is.na(flipper_length_mm)) |>
  summarize(max_flipper = max(flipper_length_mm),
            min_flipper = min(flipper_length_mm),
            count = n())
penguins |> na.omit() |>
  summarize(max_flipper = max(flipper_length_mm),
            min_flipper = min(flipper_length_mm),
            count = n())

```

* Find the max and min flipper length for each species.

```{r}
penguins |> filter(!is.na(flipper_length_mm)) |>
  group_by(species, sex) |>
  summarize(max_flipper = max(flipper_length_mm),
            median_flipper = median(flipper_length_mm),
            min_flipper = min(flipper_length_mm),
            count = n())
```


* Find the observation with the largest flipper length.

```{r}
penguins |> summarize( max_flipper_length_mm = max(flipper_length_mm) )
penguins |> summarize( max_flipper_length_mm = max(flipper_length_mm, na.rm=TRUE) )
penguins |> filter( flipper_length_mm == max(flipper_length_mm, na.rm=TRUE) )
```

  * For each combination of species and island.

```{r}
penguins |>
  group_by(species, island) |> 
  filter( flipper_length_mm == max(flipper_length_mm, na.rm=TRUE) )
penguins |>
  group_by(species) |> 
  filter( flipper_length_mm == max(flipper_length_mm, na.rm=TRUE) )
penguins |>
  group_by(island) |> 
  filter( flipper_length_mm == max(flipper_length_mm, na.rm=TRUE) )
```


* Make a table showing the number of observations of each combination of species and island.

```{r}
penguins |> 
  count(island, species)

```

We will talk next week about shifting this "long" format table into a "wide" format table that has one row per island and one column per species.

* Compute the ratio of the bill length to bill depth and make a histogram with these data.

```{r}
penguins |> 
  mutate(ratio = bill_length_mm / bill_depth_mm) |>
  # select(species, island, bill_length_mm, bill_depth_mm, ratio) |>
  ggplot(aes(x = ratio)) +
  geom_histogram(bins = 25)
```

  * For each species

```{r}
penguins |> 
  mutate(ratio = bill_length_mm / bill_depth_mm) |>
  # select(species, island, bill_length_mm, bill_depth_mm, ratio) |>
  ggplot(aes(x = ratio, fill = species)) +
  geom_histogram(bins = 25)
```

```{r}
penguins |> 
  mutate(ratio = bill_length_mm / bill_depth_mm) |>
  # select(species, island, bill_length_mm, bill_depth_mm, ratio) |>
  ggplot(aes(x = ratio, fill = island)) +
  geom_histogram(bins = 25)
```

* Compute the mean and standard deviation of body mass for each species and sex.

```{r}
penguins |>
  filter(!is.na(body_mass_g)) |>
  group_by(species, sex) |>
  summarize( mean_body_mass = mean(body_mass_g),
             sd_body_mass = sd(body_mass_g),
             n = n(),
             se_mean_body_mass = sd_body_mass / sqrt(n),
             cv_body_mass = sd_body_mass / mean_body_mass)

```


* Decode this complex calculation from the course notes by executing just one line at a time, making a prediction about the result, and checking your understanding. Change the calculation to make it do something similar but different to test your understanding. (For example, try using `slice_tail` instead of `slice_head`, or remove the `-` in the `arrange` function.)

```{r}
diamonds |>
  mutate(price_per_carat = price / carat) |>
  group_by(color, clarity, cut) |>
  summarise(median_price_per_carat = median(price_per_carat), 
            n = n(),
            .groups = "drop") |>
  arrange(-median_price_per_carat) |>
  group_by(cut) |>
  slice_head(n=2) |>
  # filter(median_price_per_carat == max(median_price_per_carat)) |>
  ungroup() |>
  arrange(-median_price_per_carat)
```


## Solutions from a previous year

How many kinds of species of penguin are in the dataset `penguins`? How many observations of each species?

```{r}
penguins |> group_by(species) |> summarize(n = n()) # These 3 lines all do the same thing
penguins |> group_by(species) |> count()
count_table <- penguins |> count(species)
penguins |> count(species)
```

What is the maximum and minimum flipper length?

```{r}
penguins |> 
  na.omit() |>
  summarize(min_flipper_length = min(flipper_length_mm),
            max_flipper_length = max(flipper_length_mm))
```

Another way of omitting missing data.

```{r}
penguins |> 
  filter(!is.na(flipper_length_mm)) |>
  summarize(min_flipper_length = min(flipper_length_mm),
            max_flipper_length = max(flipper_length_mm))
```

Do this for each species:

```{r}
penguins |> 
  filter(!is.na(flipper_length_mm)) |>
  group_by(species) |>
  summarize(min_flipper_length = min(flipper_length_mm),
            max_flipper_length = max(flipper_length_mm))
```


Find the observation with the largest flipper length.

```{r}
penguins |>
  filter(!is.na(flipper_length_mm)) |>
  group_by(species, sex) |>
  mutate(max_flipper_length = max(flipper_length_mm)) |>
  filter(flipper_length_mm == max_flipper_length) |>
  select(species, sex, flipper_length_mm, max_flipper_length)
```

Combine two steps

```{r}
penguins |>
  filter(!is.na(flipper_length_mm)) |>
  group_by(species, sex) |>
  filter(flipper_length_mm == max(flipper_length_mm)) |>
  select(species, sex, flipper_length_mm)
```

Make a table showing the number of observations of each combination of species and island.

```{r}
penguins |> count(species, island)
penguins |> group_by(species, island) |> summarize(n = n())
```

Follow up with a second summarize(n) to see what happens when the island grouping is dropped.
Answers the question: How many islands was each species found on?

```{r}
penguins |> group_by(species, island) |> summarize(n = n()) |> summarize(n = n())
```

Compute the ratio of the bill length to bill depth and make a histogram with these data.

```{r}
penguins |> mutate(bill_ratio = bill_length_mm / bill_depth_mm) |>
  filter(!is.na(bill_ratio)) |>
  ggplot(aes(x = bill_ratio)) +
  geom_histogram(binwidth = 0.1)
```


Repeat that calculation for each species.

```{r}
penguins |> 
  mutate(bill_ratio = bill_length_mm / bill_depth_mm) |>
  filter(!is.na(bill_ratio)) |>
  ggplot(aes(x = bill_ratio, fill = species)) +
  geom_histogram(binwidth = 0.1)

```

Compute the mean and standard deviation of body mass for each species and sex.

```{r}

```


