---
title: "Working with models"
author: "Andrew Irwin"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
library(gapminder)
library(tidyverse)
library(lubridate)
```


In this demonstration, we will practice adding smooths (regression lines) to plots, including using colour and facets to create multiple lines.

## Life expectancy by year

Build up a complex plot piece by piece.

```{r}
gapminder |> ggplot(aes(year, lifeExp)) + geom_point()
gapminder |> ggplot(aes(year, lifeExp)) + geom_point() + 
  facet_grid(continent ~ .)
gapminder |> ggplot(aes(year, lifeExp)) + 
  geom_smooth() + 
  facet_grid(continent ~ .)
gapminder |> ggplot(aes(year, lifeExp)) + 
  geom_smooth(method = "lm") + 
  facet_grid(continent ~ .)
gapminder |> ggplot(aes(year, lifeExp)) + 
  geom_smooth(aes(group = country), se = FALSE, color = "#00000040", size = 0.4) + 
  geom_smooth(color = "blue") + 
  facet_grid(continent ~ .) + 
  theme_bw()

```

Quantile regression

```{r}
gapminder |> ggplot(aes(year, lifeExp)) + 
  geom_quantile() + 
  facet_grid(continent ~ .)
gapminder |> ggplot(aes(year, lifeExp)) + 
  geom_quantile(quantiles = c(0.1, 0.5, 0.9), formula = y ~ poly(x,2)) + 
  facet_grid(continent ~ .)
```

## CO2

Get data from previous lesson (reading data):

```{r}
co2 <- read_table("https://www.esrl.noaa.gov/gmd/webdata/ccgg/trends/co2/co2_mm_mlo.txt",
                  comment = "#", col_names = FALSE)
co2 |> ggplot(aes(X3, X4)) + geom_point(size = 0.4) + 
  labs(x = "Year", y = "Atmospheric pCO2 (ppmv)")
co2 |> ggplot(aes(X3, X4)) + geom_point(size = 0.4) +
  geom_smooth() + 
  labs(x = "Year", y = "Atmospheric pCO2 (ppmv)")
co2 |> ggplot(aes(X3, X4)) +
  geom_smooth() + 
  labs(x = "Year", y = "Atmospheric pCO2 (ppmv)")
```

What's the annual cycle?
Subtract the mean CO2 computed from each year. Compute the fractional part of th e year.

```{r}
co2 |> group_by(X1) |>
  mutate(fractional_year = X3 - X1, 
         mean_co2 = mean(X4),
         co2_deviation = X4 - mean_co2) |>
  ungroup() |>
  ggplot(aes(fractional_year, co2_deviation)) + 
  geom_point()
```


```{r}
co2_modified <- co2 |> group_by(X1) |>
  mutate(fractional_year = X3 - X1, 
         mean_co2 = mean(X4),
         co2_deviation = X4 - mean_co2) |>
  ungroup() 
co2_modified |>
  ggplot(aes(factor(X2), co2_deviation)) + 
  geom_boxplot()
co2_modified |>
  ggplot(aes(fractional_year, co2_deviation)) + 
  geom_smooth() + theme_bw()
lubridate::date_decimal(2022.10)
lubridate::date_decimal(2022.77)
lubridate::decimal_date(ymd("2022-10-11"))
```

Make a decade by decade plot

```{r}
co2 |> mutate(decade = 10*floor(X1/10),
              decimal_decade = X3 - decade) |>
  ggplot(aes(decimal_decade, X4)) + 
  geom_point() + 
  facet_grid(decade ~ . , scale = "free_y")
```


```{r}
co2 |> mutate(decade = 10*floor(X1/10),
              decimal_decade = X3 - decade) |>
  ggplot(aes(decimal_decade, X4)) + 
  geom_smooth(method = "loess", span = 0.1) + 
  geom_point(size = 0.2) + 
  facet_grid(decade ~ . , scale = "free_y")
```

```{r}
co2 |> mutate(decade = 10*floor(X1/10),
              decimal_decade = X3 - decade) |>
  ggplot(aes(decimal_decade, X4)) + 
  geom_quantile(quantiles = c(0.1, 0.9)) + 
  geom_point(size = 0.2) + 
  facet_grid(decade ~ . , scale = "free_y")
```

## GAM is just linear regression

```{r}
co2 |> filter(X1 > 2019) |>
  ggplot(aes(X3, X4)) + 
  geom_smooth(method = "gam", formula = y ~ s(x, bs = "cs")) +
  geom_point()
```
