---
title: "Reports"
author: "Andrew Irwin"
date: "Thursday 29 February, 2024"
output:
  html_document: default
---

```{r setup, include=FALSE}
library(tidyverse)
```

In this live demonstration I'll review many important features of R markdown reports and give tips for how to work productively while writing reports you can reproduce later.

```{r}
library(palmerpenguins)
peng_subset <- penguins |> filter(species == "Adelie")
peng_subset |> ggplot(aes(bill_depth_mm)) + geom_histogram()
```

## Templates

-   From course: tasks, assignments, projects
-   R studio standard template
-   Quarto alternative (need to maintain compatibility for Rmd users, need for innovation and improvements)

## Parts of an R markdown document

-   header material (between three minus signs)
-   text
-   R code

## Formatting

-   Headings (hash marks)
-   Paragraphs
-   Lists (numbered, bullets)
-   **bold**, *italics*
-   hyperlinks
-   spell check
-   outline menu (far right)
-   code block menu (bottom left)
-   more bullets.

```{r penguin_plot, warning=FALSE, fig.cap = "A histogram of bill length from three species of penguins on three islands in the Antarctic peninsula near Palmer Station."}
penguins |> ggplot(aes(bill_length_mm)) + geom_histogram()
```

## Visual mode

Contrast with 'source' mode

```{r}

```

## Code blocks or chunks

-   names
-   options: fig.cap="", echo=FALSE, message=FALSE and many more
-   [code chunk option tutorial](https://rmarkdown.rstudio.com/lesson-3.html) and [reference](https://bookdown.org/yihui/rmarkdown-cookbook/chunk-options.html)

## Computation model

-   Work as you go and develop
    -   Play button, command/control enter, typing in console
-   Knitting to ensure everything "works"
-   Output formats (HTML, MS Word, PDF, slides)

## Common mistakes

-   forget a library function call (`library(tidyverse)`)
-   read data from a path that exists only on your computer (projects and `here::here()` function)
-   Computations out of sequence
-   Leave R errors in your Rmd document

```{r}
# read_csv("test-data.csv")
# read_csv("/Users/airwin/Dropbox/test-data.csv")
# read_csv(here::here("data-viz/tasks", "test-data.csv"))
```

## How Rstudio tries to help

-   Red X sign
-   underlines
-   Environment viewer

```{r}
model1 <- lm(body_mass_g ~ species, 
             data = penguins)  # delete bracket to see red X
summary(model1)
p1 <- penguins |> ggplot(aes(body_mass_g)) + geom_histogram()
```

## Concepts to reinforce and reexplain

-   Objects come in different types
    -   numbers, text, dates
    -   tables
    -   plots
    -   models (of various kinds)
    -   many more
-   How to learn about the object you've created
    -   type name in console
    -   look in environment tab
-   Chain commands together with pipes (`|>` or `%>%`) vs. assigning objects to temporary names to break down computations into component parts

Do a calculation two ways: with just pipes, and then in a series of steps with named objects I'll create.

```{r}
penguins |> na.omit() |>
  group_by(species) |>
  summarize(mean_body_mass = mean(body_mass_g))
```

"step-by-step"

```{r}
penguins_no_na <- penguins |> na.omit()
penguins_group <- penguins_no_na |> group_by(species)
penguins_group |> summarize(mean_body_mass = mean(body_mass_g))
```
