---
title: "PCA demo"
author: "Andrew Irwin"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggfortify) # autoplot function
library(broom) # tidy function
```


Demonstration of a PCA.

1. Start with a dataset with several quantitative variables
2. Standardize each variable to have mean 0 and standard deviation 1, or use "correlation" option below
3. Compute PCA
4. Visualize PCA as biplot, screeplot

Try the `midwest` data:

```{r}
glimpse(midwest)
```

Make a subset with only quantitative data.

```{r}
midwest_ss <- midwest |> slice_sample(n = 100) |>
  select(-category, -PID, -county, -state)
midwest_pca <- prcomp(midwest_ss, 
                      center = TRUE,
                      scale. = TRUE)
```


Get a text summary of the PCA results showing the standard deviation associated with each principal component and the proportion of variance for each PC.

```{r}
summary(midwest_pca)
```


Show a biplot using ggplot (from ggfortify package)

```{r}
autoplot(midwest_pca, data = midwest_ss, loadings = TRUE, 
         loadings.label = TRUE, label = FALSE, loadings.label.size = 3,
         variance_percentage = TRUE)
```

Show the row number of each observation instead of a dot:

```{r}
autoplot(midwest_pca, data = midwest_ss, 
         shape = FALSE, label = TRUE,
         loadings = TRUE, loadings.label = TRUE, 
         variance_percentage = TRUE)
```

What is row 44?

```{r}
midwest_ss[44,]

```

Show a similar plot using a different function:

```{r}
biplot(midwest_pca)
```

Show the variances associated with each PC.

```{r}
screeplot(midwest_pca)
```

Show the vector of each PC in terms of the orginal variables:

```{r}
tidy(midwest_pca, "loadings") |> pivot_wider(names_from = PC, values_from = value, names_prefix = "PC ")
```

Alternative function to do the same computation.

```{r}
midwest_pca2 <- princomp(midwest_ss, cor = TRUE)
summary(midwest_pca2)
biplot(midwest_pca2)
```

ggplot plots

```{r}
autoplot(midwest_pca2, data = midwest_ss, loadings = TRUE,
         loadings.label=TRUE)

```

## Scaling experiment

Don't scale the variables. You rarely really want to do this!

```{r}
midwest_ss <- midwest |> slice_sample(n = 100) |>
  select(-category, -PID, -county, -state)
midwest_pca3 <- prcomp(midwest_ss, 
                      center = TRUE,
                      scale. = FALSE)
autoplot(midwest_pca3)
screeplot(midwest_pca3)
tidy(midwest_pca3, "loadings") |> pivot_wider(names_from = PC, values_from = value, names_prefix = "PC ")

```


## Same thing with penguins

```{r}
library(palmerpenguins)
my_penguins_raw = penguins_raw |> select(-`Sample Number`) |>
  select(Species, where(is.numeric) ) |> na.omit() |>
  mutate(Species = str_remove(Species, "\\(.*\\)"))
pca2 <- my_penguins_raw |> select(-Species) |> prcomp(scale=TRUE)
autoplot(pca2, data = my_penguins_raw, 
         loadings=TRUE, loadings.label = TRUE,
         loadings.label.colour = "black", loadings.colour = "black",
         colour = 'Species') + xlim(-0.15, 0.15) + ylim(-0.15, 0.15)
```


```{r}
pca2a <- my_penguins_raw |> select(-Species) |> prcomp(scale=FALSE)
autoplot(pca2a, data = my_penguins_raw, 
         loadings=TRUE, loadings.label = TRUE,
         loadings.label.colour = "black", loadings.colour = "black",
         colour = 'Species') + xlim(-0.15, 0.15) + ylim(-0.15, 0.15)
```
