---
title: "Vector maps"
author: "AJ Irwin"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Vector maps

* Packages
* Key functions
* Data for maps
* Coloring regions
* Adding points
* Projections

## Packages

```{r}
library(tidyverse)
library(ggmap)
library(rnaturalearth)
library(rnaturalearthdata)
library(rnaturalearthhires)
```

Lots of other packages in examples (notes and wider internet): `sf` (simple features), `rgdal` (geographic data abstraction library), and many more.

## Functions

* `map_data` (get data for a map, part of ggplot2)
* `geom_map` (draw a map)
  * aesthetics: `map_id` (column with names for each region to be plotted), required argument: `map` (a data frame with the map data)
* `expand_limits` (if part of your map is not showing)
* `coord_map` (projections, mercator is default)

## Data for map

```{r}
m1 <- map_data("world")
# france, italy, lakes, nz, county (USA), state (USA), usa (coastline), world, world2 (centered on Pacific Ocean)
# canada.cities, world.cities
# see: help(package = "maps") for details
```

There are many "regions" and "subregions" within some of these maps:

```{r}
count(m1, region) |> head()
m1 |> filter(region == "Canada") |> head()
```

Political borders and territorial disputes: maps will undoubtedly be incomplete, inaccurate, or conflict with some people's political views whether in Canada (Québec, Labrador), France (St Pierre & Miquelon), China, Palestine, Russia, etc.

## Make a map

```{r}
ggplot(m1, aes(map_id = region, x = long, y = lat, group=group)) +
    geom_map(fill = "white", color = "black", size = 0.5, map = m1) 
```

Show just one region

```{r}
m2 <- m1 |> filter(region == "Canada") 
ggplot(m2, aes(map_id = region, x = long, y = lat)) +
  geom_map(fill = "white", color = "black", size = 0.5, map = m2) +
  coord_map("albers", 60, 90)
```

```{r}
m2 <- m1 |> filter(region == "Ukraine") 
m2 <- m1 |> filter(region == "Iran")
m2 <- m1 |> filter(region == "Chile")
m2 <- m1 |> filter(region == "China")
m2 <- m1 |> filter(region == "France") # m2 <- map_data("france") # more detail
ggplot(m2, aes(map_id = region, x = long, y = lat)) +
  geom_map(fill = "white", color = "black", size = 0.5, map = m2) +
  coord_map("mollweide")
```

```{r}
m2 <- map_data("world2")
ggplot(m2, aes(map_id = region, x = long, y = lat)) +
  geom_map(fill = "white", color = "black", size = 0.5, map = m2) +
  # coord_map("gilbert") 
  coord_map("mollweide")
  # coord_map("lambert", -25, 25) # not good for the whole globe
  # coord_map("albers", -25, 25)

```

Get a list of projections from help("mapproject") (in `mapproj` package).



## Colouring regions

```{r}
values <- tibble( region = m1 %>% pull(region) %>% unique())
values <- values %>% mutate(value = runif(nrow(values)))
map2 <- ggplot(values, aes(map_id = region)) + 
        geom_map(aes(fill = value),
                 map = m1,
                 colour = "white", 
                 alpha = 1, size=0.2) +
   expand_limits(x = c(-180,180), y = c(-80,80)) +
   theme_bw() +
  coord_map("mollweide")
map2
map2 + theme_void() # removes axes, labels; leaves legend
```

Show just a few regions

```{r}
map2 <- ggplot(m1, aes(map_id = region)) + 
  geom_map(fill = "gray", color = "white", map = m1) + 
  geom_map(aes(fill = value),
           data = values |> slice_sample(n = 50),
           map = m1,
           colour = "white", alpha = 1, size=0.2) +
   expand_limits(x = c(-180,180), y = c(-80,80)) +
   theme_bw() +
  coord_map("mollweide")
map2 + theme_void()
```


## Adding points

Start with some data:

```{r}
library(maps)
canada.cities |> head()
```

Add points to a map:

```{r}
library(ggrepel)
m2 <- m1 |> filter(region == "Canada") 
ggplot(m2) +
  geom_map(aes(map_id = region),
           fill = "white", color = "black", size = 0.5, map = m2) +
  geom_point(data = canada.cities,
             aes(x = long, y = lat),
             size = 0.25) +
  geom_label_repel(data = canada.cities |> filter(pop > 500000),
             aes(x = long, y = lat, label = name),
             size = 3) +
  geom_point(data = canada.cities |> filter(pop > 500000),
             aes(x = long, y = lat),
             size = 0.25, color = "red") +
  coord_map("albers", 60, 90)
```

