---
title: "Factors Dates"
author: "AJ Irwin"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
```

## Text

### Cleaning up input, for example for data entered in a spreadsheet.

```{r}
my_string <- "   A cat  is a small 
                    and furry animal.  "
my_string2 <- str_squish(my_string)
my_string2
str_to_lower(my_string2)
str_to_upper(my_string2)
str_to_sentence(my_string2)
str_to_title(my_string2)
```


### Converting text to numbers

```{r}
text_and_numbers <- tibble( text = c("Andrew", "33", "12.45", 
                                     "-1.00", "Inf"))
text_and_numbers |> mutate(numbers = as.numeric(text), 
                            integers = as.integer(text)) |> kable()
```

### Extracting information from longer labels

```{r}
sets <- c("A1", "A2", "B1", "B4", "C5")
str_extract(sets, "[0-9]")
str_extract(sets, "[A-Z]")
```

### Inserting and extrating data from longer messages

```{r}
library(glue)
library(unglue)
a <- 1
b <- 6
c <- 15.63
my_string3 <- glue("The numbers a, b, and c are {a}, {b}, and {c}, respectively. Their sum is {a+b+c}.")
my_string3
unglue(my_string3, "The numbers a, b, and c are {a}, {b}, and {c}, respectively. Their sum is {d}.")
my_strings1 <- tibble(greeting = c("My name is Andrew.", 
                                   "My name is Li.", 
                                   "My name is Emily."))
unglue_unnest(my_strings1, 
              greeting, 
              "My name is {name}.", 
              remove=FALSE) |> kable()
```


## Plots and text labels (factors)

```{r}
mpg |> ggplot(aes(x = cty,
                   y = trans)) +
  geom_boxplot()
```

```{r}
mpg |> ggplot(aes(x = cty,
                   y = fct_reorder(trans, cty, min, .desc=TRUE))) +
  geom_boxplot() 
```

Simplify the number of categories

```{r}
mpg |> 
  ggplot(aes(y = cty,
             x = fct_lump(trans, 4) |> fct_reorder(cty, median))) +
  geom_boxplot() 
mpg |> 
  ggplot(aes(x = cty,
             y = fct_lump(trans, 4) |> fct_reorder(cty, median))) +
  geom_boxplot() 
```


## Dates and times

Using lubridate package

```{r}
today()
now()
now(tz = "UTC")
now(tz = "America/Toronto")
now(tz = "Asia/Shanghai")
```

List of timezones: https://en.wikipedia.org/wiki/List_of_tz_database_time_zones

### Text to date

```{r}
dt1 <- tibble(text_date = c("1999-01-31", "2000-02-28", "2010-06-28",
                            "2024-03-14", "2021-02-29"),
             date = ymd(text_date))
dt1 |> arrange(date)
```

Crazy formats!

```{r}
tibble(date = c("Jan 5, 1999", "Saturday May 16, 70", "8-8-88",
               "December 31/99", "Jan 1, 01"),
      decoded = mdy(date)) |> kable()
```

With times


```{r}
dt2 <- tibble(text_date = c("1999-01-31 09:14", "2000-02-28 12:15",
                            "2010-06-28 23:45", 
                            "2023-03-10 00:15", "2023-03-10 01:15", "2023-03-10 02:15", "2023-03-10 03:15",
                            "2024-03-14 07:00 AM", "2021-03-01 6:16 PM"),
             date_time = ymd_hm(text_date, tz="America/Halifax"))
dt2 |> kable()
```

### Converting dates to numbers

```{r}
t1 <- now()
year(t1)
day(t1)
hour(t1)
decimal_date(t1)
yday(t1)
date_decimal(2022.95)
```

## Plotting with dates and times

```{r}
dt2 |> mutate(r = rnorm(n(), 20, 3)) |>
  ggplot(aes(x = date_time, y = r)) + 
  geom_point()  +
  # scale_x_datetime(date_labels = "%Y %Z\n%b-%d %H:%M:%S")
  scale_x_datetime(date_labels = "%Y-%b-%d")
```
