---
title: "Theme"
author: "Andrew Irwin"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse, quietly = TRUE)
library(palmerpenguins)
```

## Make a reference figure

```{r}
p1 <- penguins |> na.omit() |>
  ggplot(aes(body_mass_g, 
             flipper_length_mm,
             color = species,
             shape = island)) +
  geom_point()
p1
```

## Built-in themes

Try some of: 

* theme_bw, theme_void, theme_dark, theme_classic

```{r}
p1 + theme_bw()
```

Try themes from the ggthemes package
* theme_wsj, theme_tufte, theme_economist, theme_fivethirtyeight, theme_clean

```{r}
library(ggthemes)
p1 + theme_wsj()
```

## Text size

```{r}
p1 + theme(text = element_text(size = 20))
p1 + theme(legend.text = element_text(color = "green"),
           legend.title = element_text(color = "blue"))
```

Read  (some of the) help page for theme! At least enough to see what is there for future reference.

## Change some of the "boxes" in the theme

```{r}
p1 + theme(plot.background = element_rect(color = "orange",
                                          fill = "yellow"),
           legend.background = element_rect(color = "black",
                                            size = 0.2,
                                            linetype = "dotted"),
           panel.background = element_rect(color = "blue",
                                           fill = "#FF88FF"),
           panel.grid = element_blank(),
           axis.ticks.length = unit(1, "cm"))
```

## Change the aspect ratio

```{r}
p1 + theme(aspect.ratio = 0.4)
```

