---
title: "Graphics output"
author: "Andrew Irwin"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse, quietly = TRUE)
library(palmerpenguins)
```


## Create a figure

```{r}
p1 <- penguins |> na.omit() |>
  ggplot(aes(bill_depth_mm, 
             bill_length_mm,
             color = species,
             shape = island)) +
  geom_point()
p1
```

## Create files in different formats and sizes

Focus on creating files separate from the Rmd or HTML output.

### Bitmapped images

```{r}
ggsave("penguins-2022-11-24.png", p1,
       dpi = 300, 
       width = 16, height = 10, units = "cm")
ggsave("penguins-small-2022-11-24.png", p1,
       dpi = 300, 
       width = 8, height = 5, units = "cm")
ggsave("penguins-small-2022-11-24.jpg", p1,
       dpi = 100, 
       width = 10, height = 6, units = "cm")
```


### Vector images

```{r}
ggsave("penguins-2022-11-24.pdf", p1,
       width = 16, height = 10, units = "cm")
ggsave("penguins-small-2022-11-24.pdf", p1,
       width = 8, height = 5, units = "cm")
ggsave("penguins-small-2022-11-24.svg", p1,
       width = 10, height = 6, units = "cm") # open with web browser
```

## Formatting graphics in the Rmd output


Normal figure, not altered in any way.

```{r}
p1
```

Custom height (4 inches), width (8 inches). (What does this mean on the screen?)

```{r fig.height=4, fig.width=8}
p1
```

You won't see a change in Rstudio, but you will see a change in the webpage (HTML) output. 

Output width 50% of "page", centred left to right.

```{r out.width='50%', fig.align='center'}
p1
```

SVG output (size changes when you zoom in the browser), Aspect ratio 1.4, width = 5 (inches)

```{r fig.asp=1.4, fig.width = 5, dev = 'svglite'}
p1
```

## New formats

"WebP" as a replacement for png and jpeg.

```{r}
library(webp)
library(png)
write_webp(readPNG("penguins-2022-11-24.png"), "penguins-2022-11-24.webp")
```
