---
title: "Task: Map alternatives"
author: "Your name"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(statebins)
library(covdata)
```

In this task you will practice showing geographic data, but not on a map.

Kieran Healy (author of our main textbook) has been doing some work collecting and organizing COVID data. He writes up some of his explorations in a [blog](https://kieranhealy.org/blog/archives/2020/03/21/covid-19-tracking/) and he has collected some of this data in a R package called [covdata](https://kjhealy.github.io/covdata/). He has prepared a [codebook](https://kjhealy.github.io/covdata/articles/codebook.html) to describe the data in the package.

Install the covdata package using 

```
remotes::install_github("kjhealy/covdata@main")
```

In early September there was a problem with this package. If this problem persists when it is time to work
on this Task, I'll send some new instructions.

## Question 1

Use `covnat_daily` which reports data from European Centres for Disease Control.

Here is a list of countries.

```{r}
countries <- c("Austria", "Belgium", "Bulgaria", "Canary Islands", "Cyprus", "Estonia", "Finland", "Germany", 
               "Greece", "France", "Italy", "Spain", "Portugal", "Ireland", "Macedonia", "Luxembourg", 
               "UK", "Switzerland", "Norway", "Sweden")
```

Select data from `covnat_weekly` that match these countries or regions (some may be missing, that's OK) with date after 2020-12-01 using filter and the comparison operator `%in%`. Use ggplot and geom_tile to show cases (as fill colour) for those dates on one axis and the country on the other axis.

```{r}


```

Make a second heatmap to show the per captia case rate: cases / pop.

```{r}


```


## Question 2

Show data from the data frame `covus` using `geom_statebins`. Pick a date (`date`) and a quantitative variable (described by `measure` and reported in `count`). Use `filter` to select those data for each state. Make sure you have only a single number (or NA) for each state. Show the data on a "statebins" map.

There may be some state codes that don't match the map used by statebins. That's okay.

```{r}


```



## Submit?

If you want to submit this task for evaluation, change the FALSE to TRUE in the following code block. Save and knit your R markdown document by clicking Knit in the toolbar at the top of the editing window. If there are errors when you knit the document you will not receive full marks.

```{r}
assess_task = FALSE
```

