---
title: "Task: Summarizing data"
author: "Your Name"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(gapminder)
```

The purpose of this task is to practice the use of `filter`, `mutate`, `group_by` and `summarize` from the `dplyr` package. This is Task 5 for Lesson 9 on Summarizing Data.

For each question you should filter, transform or summarize the data, make a visualization of your data, and answer a question using the calculation and visualization with a written sentence.

## Question 1

Filter the gapminder data keeping only observations from Canada (use `filter(country == "Canada")`). (I've done this first step for you.) Then make a line plot (`geom_line`) of life expectancy as a function of time.

```{r}
#| fig-cap: "Change this text to be a meaningful caption for your figure."
gapminder |> filter(country == "Canada")
```

By approximately how many years has the life expectancy increased in Canada over the time described by this data?

**Replace this text with your answer written in a complete sentence.**


## Question 2

Compute the median life expectancy for each year and continent, using `group_by` and `summarize`. Then plot median life expectancy vs year using a different colour for each continent.

```{r}
gapminder |> group_by(year, continent)
```

Which continent has the highest median life expectancy in 2007, according to this data?

**Replace this text with your answer written in a complete sentence.**

## Question 3

Compute a scaled version of the GDP per capita by dividing gdpPercap for each country by its maximum value using `mutate`. All countries will have a maximum value of 1. Then compute the median for each year and continent and plot the median as a function of time.

This calculation requires a few steps, which you then translate into R code:

* find the maximum GDP per capita for each country and add this as a column using mutate,
* compute a new variable which is GDP per capita divided by this maximum value,
* grouping by year and continent, compute the median of the scaled GDP per capita you just computed,
* make a line plot of these data, using colour to identify the continent.

```{r}
#| fig-cap: "Change this text to be a meaningful caption for your figure."
# write your code here
```

On which continent has median GDP per capita increased the most (as a proportion relative to 2007, not absolute amounts) according to this data?

**Replace this text with your answer written in a complete sentence.**

