---
title: "Delete old repositories"
format: html
editor: source
---


```{r}
library(tidyverse)
library(gh)
library(ghclass)
error("DANGEROUS! Deletes repos. Don't just render me.")
```

First task when starting a new offering of the course: delete the repositories from the last offering.

Three steps:

* ensure permissions are correct
* get list of repositories
* delete repositories

## Ensure permissions are correct

Make sure you have admin permissions for the organization. If you don't, ask the course admin to give you admin permissions.

## Get list of repositories

Get the list of repositories in the organization. You can do this using the GitHub API or by using a tool like `gh` (GitHub CLI).

```{r}
gh_whoami()
org_sitrep("Dalhousie-AndrewIrwin-Teaching")

repos <- org_repos("Dalhousie-AndrewIrwin-Teaching")
```

Assignments to delete

```{r}
del_assign <- repos[str_detect(repos, "data-viz-assignments-")]
sapply(del_assign, repo_delete, prompt=FALSE)
```

Projects to delete

Assignments to delete

```{r}
del_project <- repos[str_detect(repos, "data-viz-project-")]
sapply(del_project, repo_delete, prompt=FALSE)
```

