---
title: "Downloading and updating existing git repos"
author: "AJ Irwin"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse, quietly = TRUE)
library(ghclass)
```

## Tasks

* Download student work from github
* Update files if an error was made in the initial assignment

## Download student work

To mark student work, you need to 

* get a list of repositories
* indicate where they should go
* clone (or pull) each one


```{r}
repo_location <- "~/Dropbox/Teaching/STAT 2430 Data Viz/2024W/Student-assignments/repos"
dir.create(repo_location, recursive = TRUE)
github_whoami()
org_name = "Dalhousie-AndrewIrwin-Teaching"
org_sitrep(org_name)
# repos <- org_repos(org_name, filter = "data-viz-assignments-.*") # student assignment repositories
repos <- org_repos(org_name, filter = "data-viz-project-.*") # student assignment repositories
```

Clone them all into `repo_location`. (Unless the repo exists. Then pull instead. The file `.DS_Store` on a mac can create conflicts when pulling, if the student has committed this file.)

```{r}
for (repo in repos) {
  repo_dir <- paste0(repo_location, "/", basename(repo))
  if (!file.exists(repo_dir)) {
     local_repo_clone(repo, 
                      local_path = repo_location)
  } else {
    local_repo_pull(repo_dir)
  }
}
```

To pull changes

```{r}
repo_dirs = list.files(repo_location)
for (repo in repo_dirs) {
  local_repo_pull(repo)
}
```


## Update files

To update files you need to 

* pull any changes that have been made since you last updated the local copies of the repos (see previous section)
* make changes to the file in each repository
* stage and commit those changes
* push the changes back to the repo
* remind students to pull the changes back to their computer

Example update: remove .html from .gitignore, add it, and then push. Can be done from R but probably easiest to do from shell.

```
for DIR in *
do
   (cd $DIR; git pull; cp ../.gitignore .; git add .gitignore; git commit -m "fix error in .gitignore"; git push)
done
```


