---
title: "Task: LOESS and GAM"
author: "Your name"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(broom)
library(kableExtra)
library(quantreg)
```

In this task you will practice fitting LOESS and GAM smooths and adding them to visualizations.

## Question 1

Using the same data from the Linear Models Task, plot the median duration of unemployment as a function of the date. Add a GAM or LOESS smooth to the data which you think captures the main trend in the data.

```{r}


```

**Explain your preferred choice. (Why is GAM better than LOESS, or the reverse?) What is the main trend highlighted by your smooth?**

Write your response here.


## Question 2

In the code block below I calculate the increase in population each year, relative to lowest population in the year. The first and last year in the dataset are outliers because there is less than a full year of data for those years.

```{r}
my_econ3 <- economics |> mutate(year = year(date),
                                 month = month(date)) |>
              group_by(year) |>
              mutate(pop_incr = (pop - min(pop))/min(pop)) |>
              filter(pop_incr == max(pop_incr)) 
```

Plot the population increase (`pop_incr`) as a function of the year. Use `filter` to exclude the first and last years. Add a smooth curve to the graph. Choose the type you did not use in question 1 (e.g., if you used a GAM in question 1, use a LOESS here.)

```{r}

```

**Summarize the trend in the annual rate of population increase in the USA in a sentence or two.**

Write your response here.



## Submit?

If you want to submit this task for evaluation, change the FALSE to TRUE in the following code block. Save and knit your R markdown document by clicking Knit in the toolbar at the top of the editing window. If there are errors when you knit the document you will not receive full marks.

```{r}
assess_task = FALSE
```


