---
title: "Task: Factors and Dates"
author: "Your name"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(COVID19)
library(lubridate)
```

In this task you will practice ordering factors on plots and working with dates.

## Question 1

Make a boxplot of the `diamonds` data. Show the distribution of the quantitative data carat for each of the levels of the cut categorical variable. Order the cut categorical variable using `fct_reorder` according to the median of the price.

```{r}


```


## Question 2

The `COVID19` package provides access to data from many countries. The second variable is `date`. You can use the function `wdoy` from the lubridate package to determine the day of week for each observation (Sunday = 1).

For example,

```{r}
covid_canada <- covid19("CA", start = "2022-01-01", verbose=FALSE)
wday(today())
wday(today(), label=TRUE)
```

Add a new variable to `covid_canada` equal to the day of the week. Filter the data to only include observations with the day of the week equal to 1 (or any other day you choose). Make a bar chart (`geom_col`) of the cumulative number of cases (`confirmed`) with a bar for each week of the year in the dataset.

```{r}


```

The "epidemiological week" begins on Saturday. The function `epiweek` gives you a value from 1-53 for any date. Compute and plot the median number of people hospitalized (`hosp`) reported in each epidemiological week.

```{r}


```

**In what week after the January wave did the number of people hospitalized reach a minimum?**

## Submit?

If you want to submit this task for evaluation, change the FALSE to TRUE in the following code block. Save and knit your R markdown document by clicking Knit in the toolbar at the top of the editing window. If there are errors when you knit the document you will not receive full marks.

```{r}
assess_task = FALSE
```


