---
title: "Write a title here"
author: "Your name"
date: "`r Sys.Date()`"
format:
  revealjs:
    slide-number: true
---

```{r include=FALSE}
# load libraries here, but show no output
library(tidyverse)
library(palmerpenguins)
```

# Instructions

* Modify this file to be a short collection of slides for a presentation
* Follow the instructions on the rest of the slides
* Delete this slide of instructions
* If you get stuck, review the [overview of making presentation slides](https://quarto.org/docs/presentations/revealjs/)

## Headings

Learn, by experimentation, the difference between a heading with one (`#`) or two (`##`) hash marks.

**Describe the difference here.**

## Images

Find a small graphic (photo, drawing, etc.). Add it to this repository. Make it appear on this slide.

Follow the instructions [here](https://quarto.org/docs/authoring/figures.html).

::: footer
Learn how to put an image in the [background](https://quarto.org/docs/presentations/revealjs/#image-backgrounds).
:::

# Visualization {.smaller}

Change the ggplot graphic here to something of your choosing. 

(You can create a new one, or use something from earlier in the course.)

Create a caption for the figure. 

```{r echo=FALSE}
#| label: fig-one
#| fig-cap: "A descriptive caption."
#| fig-width: 9
palmerpenguins::penguins |> 
  filter(!is.na(flipper_length_mm)) |>
  ggplot(aes(flipper_length_mm)) +
  geom_density() +
  theme_minimal() +
  theme(text = element_text(size = 18))
```

# Another slide {background-color="gray"}

Set the background to black, or another color.


::: footer
Learn more: [Slide Backgrounds](https://quarto.org/docs/presentations/revealjs/#color-backgrounds)
:::


# View the slides in a browser

* Click: Render.
* Open the slides in a web browser. (Click Present, or find the html file and open it.)
* Hit ? for help.
* Click on the menu in the lower left to see what options are available in presentation mode.

## Two columns

:::: {.columns}

::: {.column width="30%"}
Put an image or figure in this column
:::

::: {.column width="70%"}
- Write a sentence about the graphic here.
- Adjust the column widths to make good use of space.
:::

::::


Learn how to show code and how to "fold" code [here](https://quarto.org/docs/get-started/computations/rstudio.html#code-folding).

