# STAT 2430 Data Visualization {.unnumbered}

```{r  include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(gt)
```

This is the syllabus for the Data Visualization (STAT 2430) course offered in Winter 2026 at Dalhousie University.

* Instructor: Andrew Irwin, a.irwin@dal.ca
* Meeting times: Tuesday and Thursday 4:05-5:25 pm, Life Science Centre (LSC) C242
* Office hours: Tuesday and Thursday 3:00-4:00 pm  in Chase 225, or by appointment.

## Course Description {-}

Data visualization is the art and science of turning data into readable graphics. 
We will explore how to design and create data visualizations.
This exploration will include both the principles and techniques of data visualization. 
Students will learn the value of visualization, specific techniques, and how to design and improve a data visualization.

### Prerequisites {-}

At least one MATH or STAT course at or above the 1000 level. 
No experience with R or computer programming is required, but a desire to learn about both is essential.  

## Course Materials {-}
 
Course notes:

- Andrew J Irwin (2026) [Data visualization](https://andrewirwin.github.io/data-viz-notes/).

Textbooks:

- Kieran Healy (2018) [Data Visualization: A Practical Introduction](https://socviz.co). Princeton University Press. 
- Claus O. Wilke (2019) [Fundamentals of data visualization](https://clauswilke.com/dataviz/). O'Reilly.
- Hadley Wickham, Mine Çetinkaya-Rundel, Garret Grolemund (2023) [R for data science](https://r4ds.hadley.nz/). Second edition. O'Reilly. 

The textbooks are available in printed form from bookstores and online. The online versions are free to use. You are not required to buy the books on paper. All three are excellent books with very distinct goals.
The books by Healy and Wilke focus on the design and interpretation of data visualizations. The book by Wickham _et al._ describes the use of R, particularly the set of packages known as the "tidyverse" for data visualization and the broader tasks of data science.


### Software {#syllabus_software .unnumbered}

We will use the statistical software R and RStudio and version control software git. No prior experience with
R, RStudio or git is assumed. We will take class time to learn the software.

* To download and install R go to [r-project.org](https://www.r-project.org/) and click on the link to [download R](http://cran.r-project.org/mirrors.html)
* To download and install Rstudio, go to posit.co and click on the link to [download Rstudio](https://posit.co/download/rstudio-desktop/) Desktop.
* To download and install git: 
  * on Windows, go to [git-scm.org](https://git-scm.org) and click on the link to [download a version for Windows](https://git-scm.com/download/win). Install the software in the usual way.
  * on Macintosh, use the Terminal pane in the Rstudio application or the [Terminal app](https://www.dummies.com/computers/macs/mac-operating-systems/unix-terminal-application-on-your-macbook/) and type `xcode-select --install` to download and install git.

These tools can be installed on Linux computers as well. Contact me if you have trouble. If you have a Chromebook you can use all these tools through the cloud service [posit.cloud](https://posit.cloud/). Everyone should learn to use the cloud service as a backup in case of problems with their own computer. The cloud service provides 25 hours of free usage each month. If you can't install the software on your computer and you run into limits with the free account, please contact me.

R and Rstudio are available on Dalhousie computer labs, but the git version control software must be installed following the instructions for Windows computers above. Since all your user files are erased from lab computers when you log out, this process must be repeated on each login. For this reason, I don't recommend using Dalhousie computer labs for this course, unless you are using posit.cloud.

You should install this software on your computer during the first week of the course. 
If you are unable to install the software, get in touch with me for assistance.

## Course Structure {-}

Each week will be structured around the following components:

- Reading course notes and suggested material from the textbooks.
- In class meetings which will be a mixture of lectures, demonstrations, and skills practice. Some class time will be be dedicated to computer demonstrations and exercises; these times will be identified by a file describing the demonstration (marked Demo in the detailed outline). You are encouraged to bring your computer to these classes.
- Opportunities to develop and demonstrate your knowledge in course work to be evaluated (assignments, tests, and a term project).
- Scheduled office hours. Come to office hours if you have questions about the classes, readings, or work to be submitted for grading. If you can't participate in office hours, but you have a question, please get in touch with me and arrange another time to meet.

**Follow the [detailed schedule](002-detailed-outline.qmd) for each week's plan to keep you on track.**

## Evaluation {-}

Your goal in this course is to learn the principles and skills of data visualization. Most people learn best using a combination of collaborative and independent work. The general guideline is that work you submit should be your own---your ideas, your thoughts, your choices, your code, your typing. You are encouraged to discuss your work with the instructor and students. When you help another student, be careful you don't help them so much that you inhibit their learning.

**Exercises** are designed to develop and test your understanding of the key concepts from the lessons. **Exercises are not assigned or submitted for grading, but will help you with assignments and tests.**

**Assignments** combine skills from multiple lessons into a meaningful data visualization activity. These are opportunities to apply the content of lessons in thoughtful and sometimes creative ways.
Assignments are assessments of your skills and should be done independently.

The **mid-term test and final exam** are written on paper and in person and will test your knowledge of material from exercises and assignments.
 
The **Term Project** is an opportunity to combine many of the skills learned in the course. You will explore, analyze, and present an analysis of a data set of your choice. Your proposal will select a data set and describe some of your planned analyses. The oral presentation will be a 5-10 minute overview of your data and key visualizations. You will submit a recording of your oral presentation to be graded.
The term project is designed to be done collaboratively with a partner. Your project should be the work of members of the group and not outside collaborators or internet sources.

* Assignments (30%, roughly 1 every two weeks)
* Mid-term test, 20%
* Term project, divided into two components
  * Proposal, due Tuesday 3 March, 5%
  * Oral presentation, recording due last day of classes, 15%
* Final exam, 30%
  
Assignments are due on Friday, approximately one week after they are assigned. If Friday is a holiday (Feb 6, April 3) the assignment will be due the following Monday. 

Copying work from any other sources is not allowed and will be considered an academic offense for this course. You are encouraged to *learn* from many different sources. If you make use of material outside of course materials on assignments or the project report, include references to the material and a description of what you used in a "Sources used" section at the end of your work. There is a specific lesson that describes how artificial intelligence (AI, or large language models LLMs) can be used in this course.

_Late policy._ Assignments will not be accepted after they have been sent to the teaching assistant for grading. If an accommodation is requested and granted due to illness or other unforeseeable circumstance, the value of that work will be redistributed to other components of the course.

Grades will be reported on Brightspace.


### Letter grades {-}

Your numerical grade will be converted to a 
[letter grade for the course](https://www.dal.ca/campus_life/academic-support/grades-and-student-records/grade-scale-and-definitions.html) using the Dalhousie Common Grade Scale.

First your numerical grade will be rounded up to the nearest integer, then it will be converted to a letter using this table.

```{r echo=FALSE}
grade_table <- tibble(
        letter = c("A+", "A", "A-", "B+", "B", "B-", "C+", "C", "C-", "D", "F"),
        min_grade = c(90, 85, 80, 77, 73, 70, 65, 60, 55, 50, 0),
        max_grade = c(100, min_grade[1:10]-1)
)
grade_table |> unite(`Grade range`, min_grade, max_grade, sep="-") |>
        rename(`Letter grade`=letter) |>
  pivot_wider(names_from = `Letter grade`, values_from = `Grade range`) |>
  gt() |> cols_align("center") #  |> cols_width(everything() ~ pct(120))
```

## Course Policies {-}

### Data retention policy

During the course you will create work on a third-party website (github.com). These files will be shared only with the instructor, TAs, and student team members. All files will be kept for 6 months after the end of the course. Repositories will be deleted after that time. If you wish to keep files after that date, either clone the repositories to your computer or fork them to your own GitHub account separate from the course.

## Important dates {-}

* Wednesday 7 January: Term begins
* February 16-20: Study break
* Tuesday 7 April: Last Tuesday/Thursday class

If you know you will be unable to meet certain deadlines in advance, for _bona fide_ reasons such as scheduled medical appointments or religious observances, please contact me as soon as practical.

I will not require you to provide a medical note or to use the Student Declaration of Absence form.

For due dates for evaluations, see the detailed course outline and evaluation schedule.


## Territorial Acknowledgement

The Dalhousie University Senate acknowledges that we are in [Mi’kma’ki](https://native-land.ca/maps/territories/mikmaq/),
the ancestral and
unceded territory of the Mi’kmaq People and pays respect to the Indigenous knowledges held by
the Mi’kmaq People, and to the wisdom of their Elders past and present. The Mi'kmaq People
signed Peace and Friendship Treaties with the Crown, and [section 35 of the Constitution Act,
1982](https://indigenousfoundations.arts.ubc.ca/constitution_act_1982_section_35/) 
recognizes and affirms Aboriginal and Treaty rights. 

The Dalhousie University Senate also acknowledges the histories, contributions, and legacies of
African Nova Scotians, [who have been here for over 400 years](https://ansa.novascotia.ca/community). For more information see, e.g., 
the [Black Cultural Centre for Nova Scotia](https://bccns.com/our-history/)

## University Policies and Statements {-}

This course is governed by the academic rules and regulations set forth in the University Calendar and by Senate.

Please see the supplement to the syllabus on Brightspace for important statements, policies, and resources provided by the Faculty of Science and the University.
