# Syllabus {.unnumbered}

```{r  include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(gt)
```

This is the syllabus for the Data Visualization (STAT 2430) course in Fall 2022 at Dalhousie University.

* Instructor: Andrew Irwin, a.irwin@dal.ca
* Meeting times: Tuesday and Thursday 1:00-2:20 pm Atlantic in LSC-C244 and on Collaborate
* Office hours: Wednesday XXX pm Atlantic in Chase 225 and on Collaborate

## Course Description {-}

Data visualization is the art and science of turning data into readable graphics. 
We will explore how to design and create data visualizations.
This exploration will include both the principles and techniques of data visualization. 
Students will learn the value of visualization, specific techniques and understand how to select the best visualization method.

## Course Prerequisites {-}

At least one MATH or STAT course at or above the 1000 level. 
No experience with R or computer programming is required, but a desire to learn about both is essential.  

## Course Materials {-}
 
Course notes:

- Andrew J Irwin (2022) [Data visualization](https://andrewirwin.github.io/data-visualization/).

Textbooks:

- Kieran Healy (2018) [Data Visualization: A Practical Introduction](https://socviz.co). Princeton University Press. 
- Garret Grolemund & Hadley Wickham (2017) [R for data science](https://r4ds.had.co.nz/). O'Reilly. 
- Claus O. Wilke (2019) [Fundamentals of data visualization](https://clauswilke.com/dataviz/). O'Reilly.

The textbook and supplemental textbooks are available in printed form and online. The online versions are free to use. You are not required to buy the books on paper. All three are excellent books with very distinct goals.

The books by Healy and Wilke focus on the design and interpretation of data visualizations. The book by Grolemund and Wickham describes the use of R, particularly the set of packages known as the "tidyverse" for data visualization and the broader tasks of data science.

## Course Structure {-}

Each week will be structured around the following components:

- Reading course notes and suggested material from the textbooks.
- In class meetings which will be a mixture of lectures and demonstrations.
  - Some class time will be be dedicated to computer exercises, so you should bring your computer to these classes. I will alert you to these activities in the detailed course outline and by making an annoucement in the previous class meeting.
- Opportunities to develop and demonstrate your knowledge (tasks, assignments, and a term project)
- Scheduled office hours. Come to office hours if you have questions about the classes, readings, or work to be submitted for grading.

I recorded mini-lectures for most lessons the last time this course was offered. You don't need to watch these, but if you wish additional material you can find them [here](http://bio.mmab.ca/data-visualization/).


**Follow the [detailed outline](002-detailed-outline.qmd) for each week's plan to keep you on track.**

## Evaluation {-}

_Tasks_ are opportunities to test your understanding of the key concepts from the lessons and demonstrate you have developed basic proficiency with essential computing skills. Most lessons will have a task for you to complete. You are expected to submit one task to be graded each week.

_Assignments_ combine skills from multiple lessons into a meaningful data visualization activity. These are opportunities to apply the content of lessons in thoughtful and sometimes creative ways.

The _Final Project_ is an opportunity to combine many of the skills learned in the course. You will explore, analyze, and present to a reader an analysis of a dataset of your choice. Your proposal will select a dataset and describe some of your planned analyses. The oral presentation will be a 5 minute overview of your data and key visualizations. The report will be a polished reproducible document demonstrating many of the ideas of the course using your selected dataset.

* Tasks for most lessons (30%, 1 per week, due weekly, equally weighted)
* Assignments (40%, roughly 1 every two weeks, equally weighted)
* Term project, divided into three components
  * Proposal, due Friday 12 March, 5%
  * Oral presentation, due Thursday 1 April, 10%
  * Report, due Friday 9 April, 15%

Tasks are due on Friday. Assignments are due on Wednesday. Late work will be accepted without penalty until the work is graded or solutions are posted. After that, no late work will be accepted. If an additional accommodation is requested and granted due to illness or other unforseeable circumstance, the value of that work will be redistributed to other tasks or assignments. *If you anticipate not being able to submit any component of the term project (proposal, oral presentation, or report) on time, please contact me by email.*

Grades will be reported on [Brightspace](https://dal.brightspace.com/d2l/home/232100)

## Software {#syllabus_software .unnumbered}

We will use the statistical software R and RStudio and version control software git. No prior experience with
R, RStudio or git is assumed. We will take class time to learn the software.

* To download and install R go to [r-project.org](https://www.r-project.org/) and click on the link to [download R](http://cran.r-project.org/mirrors.html)
* To download and install Rstudio, go to [Rstudio.com](https://www.rstudio.com/) and click on the link to [download Rstudio](https://rstudio.com/products/rstudio/download/)
* To download and install git: 
  * on Windows, go to [git-scm.org](https://git-scm.org) and click on the link to [download a version for Windows](https://git-scm.com/download/win)
  * on Macintosh, use the [Terminal app](https://www.dummies.com/computers/macs/mac-operating-systems/unix-terminal-application-on-your-macbook/) and type `xcode-select --install` to download and install git.

These tools can be installed on Linux computers as well. Contact me if you have trouble. If you have a Chromebook you can use all these tools through the cloud service [rstudio.cloud](https://rstudio.cloud/project/1998812). Everyone should learn to use the cloud service as a backup in case of problems with setting up the software on their own computer.

R and Rstudio are available on Dalhousie computer labs, but the git version control software must be installed following the instructions for Windows computers above. Since all your user files are erased from lab computers when you log out, this process must be repeated on each login.

You should install this software on your computer very early in the course. 
If you are unable to install the software, get in touch with me for assistance.

## Course Policies {-}

### COVID-19 

**Please do not come on campus if you have symptoms similar to COVID-19 such as a fever or cough or if you have had a close contact with someone who has COVID-19.**

The course is intended to be delivered in person, on campus at Dalhousie. 
I hope that most of you will be physically present in the classroom for most classes.
Since you can be infected and tranmit the virus without experiencing any symptoms, I encourage everyone to wear a high-quality mask while on campus to reduce the probability of virus transmission.
I anticipate that some students will need to miss some classes following a close contact with an infected person, developing symptoms similar to COVID-19, or contracting COVID-19.
All course meetings will be available live on Brightspace and recorded for watching later.
This means that you can participate in the whole course even if you can't attend a particular class. 
You are not required to be physically present on campus for classes, tests, or a final exam.
This design is to keep you safe and to help you keep your classmates safe.

If I am unable to be on campus, the class will be delivered remotely.

### Evaluation

I will not mark or give course credit for coursework submitted after grading is complete or the solutions are posted. If you miss an assignment due to illness or other event outside your control, contact me for an accommodation. If an accommodation is granted, the value of the coursework will be shifted to other elements of the course. You may submit two late _tasks_, so I will be unlikely to offer accommodation for more than two late or missed tasks.

Your goal in this course is to learn the principles and skills of data visualization. Most people benefit from a mixture of collaborative and independent work. The general guideline is that work you submit should be your own---your ideas, your thoughts, your choices, your code, your typing. You are encouraged to discuss your work with the instructor and students. When you help another student, be careful you don't help them so much that you inhibit their learning.

_Tasks_ are primarily designed to help you learn and you are encouraged to seek assistance from classmates, but work you submit must be your own. _Assignments_ are assessments of your skills and should be done independently. The _final project_ and its components may be done in groups, but the work on the project should only be the work of members of the group.

Copying work from any other sources is not allowed and will be considered an academic offense for this course. You are encouraged to learn from many different sources. If you make use of material outside of course materials on assignments or the project report, include references to the material and a description of what you used in a "Sources used" section at the end of your work.

## Data retention policy

During the course you will create work on a third-party website (github). These files will be shared only with the instructor, TAs, and student team members. All files will be kept for 6 months after the end of the course. Repositories will be deleted after that time. If you wish to keep files after that date, either clone the repositories to your computer or fork them to your own GitHub account separate from the course.

## Important dates {-}

* Tuesday 6 September: Term begins
* Monday 10 October: Thanksgiving day, University closed
* November 7-11: Study break
* Fri 11 November: Remembrance Day
* Tue 6 December: Last Tuesday/Thursday class

If you know you will be unable to meet certain deadlines in advance, for bona fide reasons such as scheduled medical appointments or religious observances, please contact me as soon as pratical.

I will never require you to provide a medical note or to use the Student Declaration of Absence form.

For due dates for evaluations, see the detailed course outline and evaluation schedule.


## Letter grades {-}

Your numerical grade will be converted to a 
[letter grade for the course](https://www.dal.ca/campus_life/academic-support/grades-and-student-records/grade-scale-and-definitions.html) using the Dalhousie Common Grade Scale.

First your numerical grade will be rounded up to the nearest integer, then it will be converted to a letter using this table.

```{r echo=FALSE}
grade_table <- tibble(
        letter = c("A+", "A", "A-", "B+", "B", "B-", "C+", "C", "C-", "D", "F"),
        min_grade = c(90, 85, 80, 77, 73, 70, 65, 60, 55, 50, 0),
        max_grade = c(100, min_grade[1:10]-1)
)
grade_table %>% unite(`Grade range`, min_grade, max_grade, sep="-") %>%
        rename(`Letter grade`=letter) %>% gt() %>% cols_align("center")
```


## University Policies and Statements {-}

This course is governed by the academic rules and regulations set forth in the University Calendar and by Senate

*Missed or Late Academic Requirements due to Student Absence*.
Information on the university policy, including the use of the Student Declaration of Absence can be found [here](https://www.dal.ca/dept/university_secretariat/policies/academic/missed-or-late-academic-requirements-due-to-student-absence.html).

*Academic Integrity*.
At Dalhousie University, we are guided in all of our work by the values of academic integrity: honesty, trust, fairness, responsibility and respect (The Center for Academic Integrity, Duke University, 1999). As a student, you are required to demonstrate these values in all of the work you do. The University provides policies and procedures that every member of the university community is required to follow to ensure academic integrity. [More information](https://www.dal.ca/dept/university_secretariat/academic-integrity.html).

*Accessibility*.
The Advising and Access Services Centre is Dalhousie's centre of expertise for student accessibility and accommodation. The advising team works with students who request accommodation as a result of a disability, religious obligation, or any barrier related to any other characteristic protected under Human Rights legislation (Canada and Nova Scotia). [More information](https://www.dal.ca/campus_life/academic-support/accessibility.html).

*Student Code of Conduct*.
Everyone at Dalhousie is expected to treat others with dignity and respect. The Code of Student Conduct allows Dalhousie to take disciplinary action if students don’t follow this community expectation. When appropriate, violations of the code can be resolved in a reasonable and informal manner—perhaps through a restorative justice process. If an informal resolution can’t be reached, or would be inappropriate, procedures exist for formal dispute resolution. [Code](https://www.dal.ca/dept/university_secretariat/policies/student-life/code-of-student-conduct.html).

*Diversity and Inclusion - Culture of Respect*.
Every person at Dalhousie has a right to be respected and safe. We believe inclusiveness is fundamental to education. We stand for equality. Dalhousie is strengthened in our diversity. We are a respectful and inclusive community. We are committed to being a place where everyone feels welcome and supported, which is why our Strategic Direction prioritizes fostering a culture of diversity and inclusiveness. [Statement](http://www.dal.ca/cultureofrespect.html).

*Recognition of Mi’kmaq Territory*.
Dalhousie University would like to acknowledge that the University is on Traditional Mi’kmaq Territory. The Elders in Residence program provides students with access to First Nations elders for guidance, counsel and support. Visit or e-mail the Indigenous Student Centre (1321 Edward St) (elders@dal.ca). [Information](https://www.dal.ca/campus_life/communities/indigenous.html).

* [Important Dates in the Academic Year (including add/drop dates)](https://www.dal.ca/academics/important_dates.html)
* [University Grading Practices](https://www.dal.ca/dept/university_secretariat/policies/academic/grading-practices-policy.html)

## Student Resources and Support {-}

Please follow the following links for additional resources and support.

*Advising*: 
[General Advising](https://www.dal.ca/campus_life/academic-support/advising.html),
[Science Program Advisors](https://www.dal.ca/faculty/science/current-students/academic-advising.html),
[Indigenous Student Centre](https://www.dal.ca/campus_life/communities/indigenous.html),
[Black Advising Centre](https://www.dal.ca/campus_life/communities/black-student-advising.html),
[International Centre](https://www.dal.ca/campus_life/international-centre/current-students.html)

*Academic supports*: 
[Library](https://libraries.dal.ca/), 
[Writing Centre](https://www.dal.ca/campus_life/academic-support/writing-and-study-skills.html),
[Studying for Success](https://www.dal.ca/campus_life/academic-support/study-skills-and-tutoring.html),
[Copyright Office](https://libraries.dal.ca/services/copyright-office.html),
[Fair Dealing Guidelines](https://libraries.dal.ca/services/copyright-office/fair-dealing.html)

*Other supports and services*:
[Student Health & Wellness Centre](https://www.dal.ca/campus_life/health-and-wellness/services-support/student-health-and-wellness.html),
[Student Advocacy](https://dsu.ca/dsas),
[Ombudsperson](https://www.dal.ca/campus_life/safety-respect/student-rights-and-responsibilities/where-to-get-help/ombudsperson.html)

*Safety*:
[Biosafety](https://www.dal.ca/dept/safety/programs-services/biosafety.html),
[Chemical Safety](https://www.dal.ca/dept/safety/programs-services/chemical-safety.html),
[Radiation Safety](https://www.dal.ca/dept/safety/programs-services/radiation-safety.html),
[Scent‐Free Program](https://www.dal.ca/dept/safety/programs-services/occupational-safety/scent-free.html)
