---
title: "Making your first plot"
author: "Andrew Irwin, a.irwin@dal.ca"
date: "2021-09-13"
format:
  revealjs:
    slide-number: true
    theme:  default
    chalkboard: true
    auto-animate: true
    scrollable: true
    code-block-border-left: true
    code-copy: false
    code-link: true

---


```{r setup, include=FALSE}
library(tidyverse)
library(patchwork)
library(kableExtra)
```

# Use the ggplot2 library

Do this once (if you haven't done it already):

```
install.packages("tidyverse")
```

Add this line to every R markdown document:

```
install.packages("tidyverse")
```

## Get some data {.smaller}

```{r echo=TRUE}
glimpse(mtcars)
```

Also try `mtcars` or `str(mtcars)` or `View(mtcars)`.

## First plot

```{r fig0, fig.show="hide", warning=FALSE, fig.height=6, fig.width=6, echo=TRUE}
mtcars |>
  ggplot(aes(x = wt,
             y = mpg)) +
  geom_point()
```


- The _pipe_ symbol (`|>`) is function composition. 
    - `f(g(x))` can be written `x |> g |> f`.
- `aes` is a function to define **aes**thetic associations between features of your plot and variables in the dataset.
- Parts of a ggplot are added togther with `+`
- The kind of plot is called its _geometry_. 
    - `geom_point` makes a scatterplot.

---

```{r fig.align="center", echo=FALSE}
knitr::include_graphics(knitr::fig_chunk("fig0", "png"))
```


## What if you forget ... ?

```{r fig1, fig.show="hide", warning=FALSE, fig.height=6, fig.width=6, echo=TRUE}
mtcars |>
  ggplot(aes(x = wt,
             y = mpg))
```

Try "forgetting" other parts of the code to see what goes wrong.

---

```{r fig.align="center", out.height="80%", echo=FALSE}
knitr::include_graphics(knitr::fig_chunk("fig1", "png"))
```


## Add some colour

```{r fig2, fig.show="hide", warning=FALSE, fig.height=6, fig.width=8, echo=TRUE}
mtcars |>
  ggplot(aes(x = wt,
             y = mpg,
             color = factor(cyl))) + #<<
  geom_point()
```


`cyl` is a number, so I must turn it into a categorical variable (factor) to get a discrete colour scale.

Try using `cyl` instead of `factor(cyl)`.

---

```{r fig.align="center", out.height="80%", echo=FALSE}
knitr::include_graphics(knitr::fig_chunk("fig2", "png"))
```


## Make the text larger

```{r fig3, fig.show="hide", warning=FALSE, fig.width=7, fig.height=6, echo=TRUE}
#| code-line-numbers: 6
mtcars |>
  ggplot(aes(x = wt,
             y = mpg,
             color = factor(cyl))) + 
  geom_point() +
  theme(text = element_text(size = 28))
```

## Make the text larger

```{r fig.align="center", out.width="80%", out.height="80%", echo=FALSE}
knitr::include_graphics(knitr::fig_chunk("fig3", "png"))
```


## Make the symbols larger

```{r fig3b, fig.show="hide", warning=FALSE, fig.width=7, fig.height=5, echo=TRUE}
#| code-line-numbers: 5
mtcars |>
  ggplot(aes(x = wt,
             y = mpg,
             color = factor(cyl))) + 
  geom_point(size=3) + 
  theme(text = element_text(size = 28)) 
```


I'm setting the size of all the symbols, not connecting a variable to the size aesthetic. So don't use `aes`.


`geom_point` (and any other geom) _inherits_ the aesthetics from `ggplot`.

## Make the symbols larger

```{r fig.align="center", out.width="80%", out.height="80%", echo=FALSE}
knitr::include_graphics(knitr::fig_chunk("fig3b", "png"))
```



## Customize the labels

```{r fig4, fig.show="hide", warning=FALSE, fig.width=7, fig.height=5, echo=TRUE}
#| code-line-numbers: 7-11
mtcars |>
  ggplot(aes(x = wt,
             y = mpg,
             color = factor(cyl))) + 
  geom_point(size=3) +
  theme(text = element_text(size = 28)) +
  labs(x = "Car mass (x 1000 pounds)", 
       y = "Fuel consumption (mpg)", 
       color = "Number of\nCylinders", 
       title = "Car road test results", 
       caption = "from Motor Trend magazine") 
```


## Customize the labels

```{r fig.align="center", out.width="80%", out.height="80%", echo=FALSE}
knitr::include_graphics(knitr::fig_chunk("fig4", "png"))
```


## Summary

* Start with data

* Pipe (`|>`) into `ggplot`

* Define the aesthetics: `aes(x = ..., y = ..., color = ..., shape = ...)`

* Define the geometry: `geom_point` shown here, but there are [lots more](https://ggplot2.tidyverse.org/reference/index.html#geoms)

* Customize text


## Suggested reading

* Course notes: Making your first plot

* Healy. Section 2.6. Make your first figure

* R4DS. Chapter 3: Data visualization

* Lots more detail: The [ggplot2 book](https://ggplot2-book.org/index.html)

## Task and Assignment

* Try these plotting commands on your own 

* Assignment 1: Your first plotting exercises



## Datasets to experiment with

* mtcars, iris and many other well-known data in datasets package
* penguins in palmerpenguins package
* gapminder in gapminder package  (but see website too [Gapminder](https://gapminder.org))
* diamonds in ggplot2 package
* nycflights13 in dbplyr package

Use `glimpse(mtcars)` (or other dataset) and `View(mtcars)` to look at the data.

Use ?mtcars to get documentation for the dataset, or use the search in the "Help" panel.

