---
title: "Using the grammar of graphics (ggplot2)"
author: "Andrew Irwin, a.irwin@dal.ca"
date: "2022-09-22"
format:
  revealjs:
    slide-number: true
    theme:  default
    chalkboard: true
    auto-animate: true
    scrollable: true
    code-block-border-left: true
    code-copy: false
    echo: true
    code-link: true
    history: false
---


```{r setup, include=FALSE}
library(socviz)
library(tidyverse)
```

## ggplot2 examples

A series of examples to demonstrate how to use the grammar of graphics to plan and create visualizations.

Using `gss_sm` and `gss_cat` in the `socviz` package which contains a general social survey with 2867 observations (rows) and 32 variables (columns).

Basic examples using selected geometries, plus some elaborations.

## General Social Survey data, 2016 {.smaller}

A dataset containing an extract from the 2016 General Social Survey. See http://gss.norc.org/Get-Documentation for full documentation of the variables.

```{r}
glimpse(gss_sm)
```

## gss overview {.smaller}

```{r}
glimpse(gss_cat)
```


## What data do you have?

* One (or two) categorical: `geom_bar`

* One quantitative: `geom_histogram`

* One categorical and one quantitative: `geom_boxplot`

* Two quantitative: `geom_point`


## One categorical variable

```{r warning=FALSE, message=FALSE, fig.height=4.25, fig.align="center", dpi=150}
gss_sm |> ggplot(aes(y = region)) + geom_bar()
```


## Two categorical variables

```{r warning=FALSE, message=FALSE, fig.height=4.25, fig.align="center", dpi=150}
gss_sm |> ggplot(aes(y = region, fill = degree)) + geom_bar() 
```


## Two categorical variables, using area for counts

```{r warning=FALSE, message=FALSE, fig.height=4.25, fig.align="center", dpi=150}
gss_sm |> ggplot(aes(y=income16, x = happy)) + geom_count()
```


## Same data using bars

```{r warning=FALSE, message=FALSE, fig.height=4.25, fig.align="center", dpi=150}
gss_sm |> ggplot(aes(y=income16, fill = happy)) + geom_bar() 
```

## Same data, too many colors

```{r warning=FALSE, message=FALSE, fig.height=4.25, fig.align="center", dpi=150}
gss_sm |> ggplot(aes(fill=income16, x = happy)) + geom_bar() 
```

## One quantitative variable

```{r warning=FALSE, message=FALSE, fig.height=4.25, fig.align="center", dpi=150}
gss_sm |> ggplot(aes(x = age)) + geom_histogram()
```

## 1 categorical, 1 quantitative

```{r warning=FALSE, message=FALSE, fig.height=4.25, fig.align="center", dpi=150}
gss_sm |> ggplot(aes(x = age, y=factor(childs))) + 
  geom_boxplot() + 
  labs(y="Number of children") 
```


## Two quantitative variables

```{r warning=FALSE, message=FALSE, fig.height=4.25, fig.align="center", dpi=150}
gss_cat |> slice_sample(n=2000) |> 
  ggplot(aes(x = age, y= tvhours)) + 
  geom_point(size=0.2) 
```

## Two quantitative variables

```{r warning=FALSE, message=FALSE, fig.height=4.25, fig.align="center", dpi=150}
gss_cat |> slice_sample(n=2000) |> 
  ggplot(aes(x = age, y= tvhours)) + 
  geom_jitter(size=0.2) + 
  geom_smooth() 
```

## 2 quantitative, show variability in one

```{r warning=FALSE, message=FALSE, fig.height=4.25, fig.align="center", dpi=150}
gss_cat |> ggplot(aes(x = age, y = tvhours )) + 
  stat_summary(fun.data = "mean_se")
```


## Summary

* Showed variations on introductory example with 1 or 2 categorical and quantitative variables

* Demonstrated a few embelishments (colours, standard deviations, symbol area, smooths, jitter points)

* Showed some visualizations that "don't work" -- too many categories



## Further reading

* Lots more examples with different data in the course notes

* "Extras" including axis labels, text on figures, other annotations


## Task

Practice these graphs by

* reproducing them, and

* modifying them by changing variables used in aesthetic mappings.

If you would like, use other data sources described in "Data Sources" chapter in course notes.

