---
title: "Splitting graphs into facets"
author: "Andrew Irwin, a.irwin@dal.ca"
date: "2021-09-27"
format:
  revealjs:
    slide-number: true
    theme:  default
    chalkboard: true
    auto-animate: true
    scrollable: true
    code-block-border-left: true
    code-copy: false
    code-link: true
---

```{r setup, include=FALSE}
library(tidyverse)
library(palmerpenguins)
library(gapminder)
```

# Splitting a plot into facets

* Using one categorical variable

* Using two categorical variables

* Using quantitative variables

# One categorical variable

* `facet_wrap( ~ variable)`

---

```{r}
mtcars |> ggplot(aes(x =wt, y=mpg)) + geom_point()
```

---

```{r}
mtcars |> ggplot(aes(x =wt, y=mpg)) + geom_point() +
  facet_wrap(~cyl)
```

---

```{r warning=FALSE}
penguins |>
  ggplot(aes(x=body_mass_g, y= flipper_length_mm, color = sex)) +
  geom_point() + 
  facet_wrap( ~ species)
```



# Using two categorical variables

* `facet_grid( y ~ x)`
* `facet_wrap( ~ x + y)`

---

```{r}
mtcars |> ggplot(aes(x =wt, y=mpg)) + geom_point() +
  facet_grid(gear~cyl)
```

---


```{r}
mtcars |> ggplot(aes(x =wt, y=mpg)) + geom_point() +
  facet_grid(gear~cyl, 
             labeller = labeller(.cols = function(x) paste(x, " cylinders"),
                                 .rows = label_both))
```


---


```{r warning=FALSE}
penguins |>
  ggplot(aes(x=body_mass_g, y= flipper_length_mm, color = sex)) +
  geom_point() + 
  facet_wrap(~ species + island)
```

---

```{r}
gapminder |>
  ggplot(aes(x=year, y = lifeExp, group = country)) + 
  geom_line() +
  facet_grid( continent ~ . )
```

---

```{r message=FALSE}
gapminder |>
  ggplot(aes(x=year, y = lifeExp)) + 
  geom_line(aes(group = country), color = "gray40", size=0.25) +
  geom_smooth(method = "lm", color = "blue", se=FALSE) + 
  facet_grid( continent ~ . )
```

---

# Splitting on a quantitative variable

* Start by converting the quantitative variable into a cateogorical one

For example:

```{r}
penguins |> 
  mutate(bill_categories = cut_number(bill_length_mm, n = 4)) |>
  select(species, bill_length_mm, bill_categories, everything())
```

---

```{r warning=FALSE}
penguins |> 
  mutate(bill_categories = cut_number(bill_length_mm, n = 3)) |>
  ggplot(aes(x= body_mass_g, y = flipper_length_mm, color = species)) + 
  geom_point() + 
  facet_wrap( ~ bill_categories)
```

---

# Summary

* `facet_wrap()`, `facet_grid()`

* You can set the number of rows and columns with `nrow = Y` or `ncol=X`

* You can label facets with both the variable and value using `labeller = labeller(.rows = label_both)`


# Further reading

* More examples with different data in the course notes

* R4DS [Section 3.5](https://r4ds.had.co.nz/data-visualisation.html#facets)

* Wilke [Multi-panel figures](https://clauswilke.com/dataviz/multi-panel-figures.html) discussing facets and other suggestions for combining different views of data together into one figure

* Healy [Section 4.3](https://socviz.co/groupfacettx.html#facet-to-make-small-multiples)


# Task

Practice these skills by making the plots in Task 6.

