---
title: "Displaying tables"
author: "Andrew Irwin, a.irwin@dal.ca"
date: "2022-10-06"
format:
  revealjs:
    slide-number: true
    theme:  default
    chalkboard: true
    auto-animate: true
    scrollable: true
    code-block-border-left: true
    code-copy: false
    code-link: true
    history: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(kableExtra)
```

## Displaying tables

After we collect, reshape, and summarize data, we want to display it!

* Rename and align columns
* Format (bold, italic, colour)
* Add column headings
* Group rows
* Add captions

* Consistent goal: make the table easy to interpret


## Raw data

```{r echo = FALSE}
mtcars
```

## `kable` formatting

```{r warning=FALSE, message=FALSE, fig.height=4.25, fig.align="center", dpi=150}
mtcars |> kable() |>
  kable_styling(full_width = FALSE) 
```

## Scrolling to show big tables

```{r}
mtcars |> kable() |> 
  kable_styling() |> scroll_box()
```

## Paged data to show big tables

```{r}
mtcars |> rmarkdown::paged_table()
```

## Changing column headings

```{r}
mtcars |> as_tibble(rownames = "make_model") |>
  select(make_model, mpg, cyl, wt) |>
  kable(col.names = c("Make & Model", "Miles per gallon", 
                      "Cylinders", "Weight (1000 lbs)")) |>
  kable_styling(full_width = FALSE)
```


## Formatting rows & columns

```{r}
mtcars |> as_tibble(rownames = "make_model") |> select(make_model, mpg, cyl, wt) |>
  kable(col.names = c("Make & Model", "Miles per gallon", "Cylinders", "Weight (1000 lbs)"),
        align = "lccr",
        digits = 1) |>
  column_spec(1, bold = TRUE) |>
  column_spec(3, color = "blue") |>
  row_spec(2, italic = TRUE) |>
  kable_styling(full_width = FALSE, 
                font_size=24)
```


## Headers spanning multiple columns

```{r}
mtcars |> select(mpg, cyl, hp, wt) |> kable() |>
  add_header_above(c(" " = 1, "Group A" = 2, "Group B" = 2)) |>
  kable_styling(full_width = FALSE)
```

## Grouping rows

```{r}
mtcars |> select(mpg, cyl, hp, wt) |> kable() |>
  group_rows("Group 1", 1, 4) |>
  group_rows("Group 2", 5, 9) |>
  kable_styling(full_width = FALSE)
```

## Adding caption, footnote

```{r}
mtcars |> head() |> select(mpg, cyl, hp, wt) |> 
  kable(caption="Car data from Motor Trend magazine.") |>
  kable_styling(full_width = FALSE) |>
  footnote(symbol = c("Footnote 1.", 
                      "Footnote 2")) 
```

## Summary

* Several ways to customize the formatting of a table


## Further reading

* Course notes

* `kableExtra` vignette


## Task

The lesson's Task has some exercises for practicing table presentation.



