---
title: "Linear models"
author: "Andrew Irwin, a.irwin@dal.ca"
date: "2022-10-06"
format:
  revealjs:
    slide-number: true
    theme:  default
    chalkboard: true
    auto-animate: true
    scrollable: true
    code-block-border-left: true
    code-copy: false
    code-link: true
    history: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(broom)
library(equatiomatic)
library(kableExtra)
library(gapminder)
my_theme = theme_classic() + 
  theme(text = element_text(size=18))
```

##  Linear models {.smaller}

* Describing, fitting, and using models

* Linear regression

  * Straight lines
  * Polynomials
  * log transforms
  
* Quantile regression

* Accessing model coefficients 

* Calculating residuals

* Making predictions: point estimates and standard errors or confidence intervals


##  Start with a basic graph

```{r, fig.height=4.5, fig.align="center", echo = FALSE}
my_mpg <- mpg |> mutate(city_l100km = 235.214583 / cty) 
p <- my_mpg |>
   ggplot(aes(x =displ, y= city_l100km)) + geom_point() +
   labs(x = "Engine displacement (L)",
        y = "Fuel economy, city (L/100km)") + my_theme
p
```


## Fit a straight line

```{r, fig.height=5, fig.align="center"}
m1 <- lm( city_l100km ~ displ, data = my_mpg)
summary(m1)
```


## Fit a straight line

```{r}
library(broom)
glance(m1)  |> kable(digits = 2) |> kable_styling(full_width = FALSE)
tidy(m1) |> kable(digits = 2) |> kable_styling(full_width = FALSE)
```

## Graph estimated terms

```{r fig.height=5, fig.align="center"}
tidy(m1, conf.int = TRUE) |>
  ggplot(aes(y = term, x = estimate, xmin = conf.low, xmax = conf.high)) + 
  geom_pointrange() + my_theme
```

## Polynomial regression

```{r, fig.height=5, fig.align="center"}
m2 <- lm( city_l100km ~ poly(displ,2), data = my_mpg)
glance(m2) |> kable(digits = 2) |> kable_styling(full_width = FALSE)
tidy(m2) |> kable(digits = 2) |> kable_styling(full_width = FALSE)
```


##  Log transforms

```{r, fig.height=5, fig.align="center"}
gapminder |> ggplot(aes(x = gdpPercap, y= lifeExp)) + geom_point() +
  scale_x_log10() + 
   geom_smooth(method = "lm", formula = y ~ x) + my_theme
```

##  Log transforms

```{r, fig.height=5, fig.align="center"}
m3 <- gapminder |> mutate(logGDPpercap = log10(gdpPercap)) %>%
  lm( lifeExp ~ logGDPpercap, data = . ) 
tidy(m3) |> kable(digits = 2) |> kable_styling(full_width = FALSE)
glance(m3) |> kable(digits = 2) |> kable_styling(full_width = FALSE)
```

##  Alternative: two step

```{r, fig.height=5, fig.align="center"}
d3 <- gapminder |> mutate(logGDPpercap = log10(gdpPercap))
m3 <- lm( lifeExp ~ logGDPpercap, data = d3 ) 
tidy(m3) |> kable(digits = 2) |> kable_styling(full_width = FALSE)
```

## Quantile regression

```{r  message=FALSE}
library(quantreg)
m4 <- rq(city_l100km ~ displ, data = my_mpg,
                      tau = 0.50) # quantile, between 0 and 1
tidy(m4) |> kable(digits = 2) 
glance(m4) |> kable(digits = 2) 
```

## Working with models: predictions

```{r, fig.height=5, fig.align="center"}
predict(m1, se.fit = TRUE) |> kable(digits = 2)
```

## Working with models: residuals

```{r, fig.height=5, fig.align="center"}
residuals(m1) |> as_tibble() |> kable(digits = 2)
```

## Combining with original data

```{r, fig.height=5, fig.align="center"}
bind_cols(my_mpg |> select(displ, city_l100km),
          predict(m1) |> as_tibble() |> rename(predict = value),
          residuals(m1) |> as_tibble() |> rename(residual = value)) |> kable(digits = 2)
```


## Predictions from new data

```{r, fig.height=5, fig.align="center"}
new_data = tibble(displ = seq(1.6, 7.2, by = 0.5))
new_data |> mutate(prediction = predict(m1, new_data)) |> kable(digits = 2)
```

## Model uncertainties

```{r, fig.height=5, fig.align="center"}
bind_cols(my_mpg |>  select(displ, city_l100km),
          predict(m1, se.fit = TRUE) |> as_tibble() ) |>
 distinct() |> kable(digits = 2)
```

## Model confidence intervals

```{r, fig.height=5, fig.align="center"}
bind_cols(my_mpg |> select(displ, city_l100km),
          predict(m1, interval = "confidence") |> as_tibble() ) |>
  distinct() |> kable(digits = 2)
```


## Prediction uncertainties

```{r, fig.height=5, fig.align="center"}
bind_cols(new_data,
          predict(m1, new_data, interval = "prediction") |> as_tibble() ) |>
   kable(digits = 2)
```

## Can also use augment

```{r}
augment(m1, my_mpg, interval = "confidence") |>
  select(displ, city_l100km, .fitted, .lower, .upper, .resid, everything() ) |> kable(digits =2)
```

## Same methods for other model types

```{r}
augment(m2, newdata = new_data, interval = "prediction") |>
  kable(digits = 2)
```

## Same methods for other model types

```{r}
new_data2 = tibble(logGDPpercap = seq(2.3, 5.0, 0.25))
augment(m3, newdata = new_data2, 
        interval = "prediction", level = 0.90) |>
   kable(digits = 2)
```

## Same methods for other model types

```{r}
augment(m4, newdata = new_data, interval = "confidence") |> kable(digits = 2)
```

## Summary

* Fitting regression models (lines, polynomials, transformations) to means and quantiles

* Extracting model coefficients, uncertainties, errors (with `tidy`) 

* Computing predicted values (with two kinds of errors) and residuals (compared to data)

* Most models can't make predictions from NA. NAs in data can cause problems.

## Further reading

* Course notes

* Healy Chapter 6. Work with models.

## Task

Task: Fitting and using linear models.





