---
title: "GAM and LOESS models"
author: "Andrew Irwin, a.irwin@dal.ca"
date: "2024-02-15"
format:
  revealjs:
    slide-number: true
    theme:  default
    chalkboard: true
    auto-animate: true
    scrollable: true
    code-block-border-left: true
    code-copy: false
    code-link: true
    history: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(broom)
library(gratia)
library(kableExtra)
library(lubridate)
my_theme = theme_classic() + 
  theme(text = element_text(size=18))
```

##  GAM and LOESS models

* Describing, fitting, and using models

* No (simple, easily understood) model coefficients 

* Calculating residuals

* Making predictions: point estimates and standard errors or confidence intervals

##  Start with a basic graph

```{r, fig.height=4.5, fig.align="center", echo = FALSE}
hfx <- read_rds("static/L17/selected_weather_day.rds") |> filter(station_id == 50620, TC_id == "YHZ") |>
  filter(year == 2019 | year == 2018) |> filter(!is.na(mean_temp)) |>
  mutate(year_day = yday(date) + (as.numeric(year)-2018)*365)
p <- hfx |>
   ggplot(aes(x = year_day, y= mean_temp)) + geom_point() +
   labs(x = "Days since 2018-01-01",
        y = "Mean temperature (°C)",
        title = "Halifax weather from Environment Canada") + my_theme
p
```

## Fit a cubic spline

```{r, fig.height=5, fig.align="center"}
library(mgcv)
m1 <- gam( mean_temp ~ s(year_day, bs="cs"), data = hfx)
summary(m1)
```

## Fit a cubic spline

```{r}
library(broom)
glance(m1)  |> kable(digits = 2) |> kable_styling(full_width = FALSE)
tidy(m1) |> kable(digits = 2) |> kable_styling(full_width = FALSE)
```

## Find predicted values and residuals

```{r fig.height=5, fig.align="center"}
library(gratia)
hfx |> add_fitted(m1) |> select(year_day, mean_temp, .value) |>
  add_residuals(m1)  |> kable()
```

## Graph "smooth" term

```{r fig.height=5, fig.align="center"}
hfx |> add_fitted(m1) |> select(year_day, mean_temp, .value) |>
  add_residuals(m1) |>
  ggplot(aes(year_day)) +
  geom_line(aes(y=.value), color="blue", size=2) + 
  geom_point(aes(y = mean_temp)) + my_theme
```

## Graph "smooth" term

```{r fig.height=5, fig.align="center"}
confint(m1, "s(year_day)", level = 0.95) |>
  ggplot(aes(year_day)) +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.25, fill="red") +
  geom_line(aes(y=est), size=2, color="red") + 
  geom_point(aes(y = mean_temp - mean(mean_temp)), data = hfx) + 
  my_theme
```

## LOESS smooths

```{r, fig.height=5, fig.align="center"}
m2 <- loess( mean_temp ~ year_day, data = hfx)
summary(m2)
```

##  Predictions

```{r, fig.height=5, fig.align="center"}
hfx |> select(year_day, mean_temp) |>
  mutate(prediction = predict(m2),
         residuals = residuals(m2)) |> kable(digits = 2)
```

##  Predictions

```{r, fig.height=5, fig.align="center"}
augment(m2, se_fit = TRUE) |> kable(digits = 2)
```


##  Residuals

```{r, fig.height=5, fig.align="center"}
augment(m2, se_fit = TRUE) |> 
  ggplot(aes(x = .resid)) + 
  geom_histogram(binwidth = 1) + my_theme
```

##  Predictions

```{r fig.height=5, fig.align="center"}
new_data = tibble(year_day = seq(1:730))
augment(m2, newdata = new_data, se_fit = TRUE) |>
  ggplot(aes(x = year_day, y = .fitted)) + 
  geom_line(size = 2, color="blue") +
  geom_ribbon(aes(ymin = .fitted - 2*.se.fit,
                  ymax = .fitted + 2*.se.fit), alpha = 0.25, fill="blue") +
  geom_point(aes(y = mean_temp), data = hfx) + my_theme
```

## LOESS with smaller span

```{r fig.height=5, fig.align="center"}
m3 <- loess(mean_temp ~ year_day, data = hfx, 
            degree = 1, span = 0.25)
augment(m3, newdata = new_data, se_fit = TRUE) |>
  ggplot(aes(x = year_day, y = .fitted)) + 
  geom_line(size = 2, color="blue") +
  geom_ribbon(aes(ymin = .fitted - 2*.se.fit,
                  ymax = .fitted + 2*.se.fit), alpha = 0.25, fill="blue") +
  geom_point(aes(y = mean_temp), data = hfx) + my_theme
```


## Summary

* Fitting GAM and LOESS smooths

* No model coefficients to extract 

* Computing predicted values with errors and residuals (compared to data)

## Further reading

* Course notes

* Healy Chapter 6. Work with models. 



