---
title: "k-means clustering"
author: "Andrew Irwin, a.irwin@dal.ca"
date: "2024-03-07"
format:
  revealjs:
    slide-number: true
    theme:  default
    chalkboard: true
    auto-animate: true
    scrollable: true
    code-block-border-left: true
    code-copy: false
    code-link: true
    history: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
#library(GGally)
#library(ggfortify)
library(kableExtra)
library(gapminder)
library(broom)
library(plotly)
library(palmerpenguins)
# library(vegan)
my_theme = theme_classic() + 
  theme(text = element_text(size=18))
```

## Plan

* What is k-means clustering and why do we use it?

* Demonstration 


## What is k-means?

* A tool to place observations in discrete groups

* Based on a distance calculated on several variables

* You decide how many groups to use (but there is guidance)

* There are variations on the algorithm and other methods (not discussed in this course)

## Sample data

```{r}
gapminder_ss <- gapminder |>
  filter(year == 2007) |> select(-year)
gapminder_ss |> select(-country, -continent) |> arrange(-lifeExp) |>
  kable() |> kable_styling(full_width = FALSE)
```

## Distribution of data

```{r echo = FALSE, fig.align='center', out.width="70%", fig.asp = .6}
gapminder_ss |> pivot_longer(lifeExp:gdpPercap, names_to = "variable", values_to = "value") |>
  ggplot() +
  geom_density(aes(value)) + facet_wrap(variable ~ ., scales="free", ncol=1) + theme(aspect.ratio = .6)
```

## Distribution of data

```{r echo = FALSE, fig.align='center', out.width="70%", fig.asp = .6}
gapminder_ss |> mutate(logGDP = log10(gdpPercap), logPop = log10(pop)) |> select(lifeExp, logGDP, logPop, everything()) |>
  pivot_longer(lifeExp:logPop, names_to = "variable", values_to = "value") |>
  ggplot() +
  geom_density(aes(value)) + facet_wrap(variable ~ ., scales="free", ncol=1) + theme(aspect.ratio = .6)
```

## Data preprocessing

* Approximately normally distributed, same variance

```{r}
gapminder_scaled <- gapminder_ss |> 
  mutate(s_logPop = scale(log10(pop)), 
         s_logGDP = scale(log10(gdpPercap)),
         s_lifeExp = scale(lifeExp))
```
```{r echo = FALSE, fig.align='center', out.width="70%", fig.asp = .6}
gapminder_scaled |> ggplot() +
  geom_density(aes(s_logPop, color = "Population")) + 
  geom_density(aes(s_logGDP, color = "GDP")) + 
  geom_density(aes(s_lifeExp, color = "Life Expectancy")) + theme(aspect.ratio = .6)
```

## Perform k-means

```{r}
kclust1 <- kmeans(gapminder_scaled |>
                    select(s_lifeExp, s_logGDP, s_logPop),
                  centers = 5)
tidy(kclust1) |> kable()
```

## Display clustering

```{r fig.align="center", out.width="60%", fig.asp=0.8}
kclust1 |> augment(gapminder_scaled) |>
  ggplot(aes(y = lifeExp, x = log10(pop), color = .cluster, shape = continent)) + 
  geom_point(size=2) + theme(aspect.ratio = .8)
```


## How good is the clustering?

```{r echo = FALSE, out.width="60%", fig.align="center"}
kclusts2 <- 
  tibble(k = 1:9) |>
  mutate(
    kclust = map(k, ~kmeans(gapminder_scaled |> select(s_lifeExp, s_logGDP, s_logPop),
                            centers = .x)),
    tidied = map(kclust, tidy),
    glanced = map(kclust, glance),
    augmented = map(kclust, augment, gapminder_scaled)
  )
clusters2    <- kclusts2 |> unnest(cols = c(tidied))
assignments2 <- kclusts2 |> unnest(cols = c(augmented))
clusterings2 <- kclusts2 |> unnest(cols = c(glanced))
clusterings2 |> mutate(proportion_within = tot.withinss / totss) |>
  ggplot(aes(x = k, y = proportion_within)) + 
  geom_line() + geom_point() + theme_bw() +
  scale_x_continuous(breaks = 1:9) + ylim(0,1)
```

## Second example: penguins

```{r echo = FALSE, fig.align='center', out.width="70%", fig.asp = .6}
penguins |> na.omit() |>
  pivot_longer(bill_length_mm:body_mass_g) |>
  ggplot() +
  geom_density(aes(value, color = species)) + facet_wrap(~ name, scales= "free") + 
  theme(aspect.ratio = .6)
```


## Second example: penguins

```{r}
penguins_scaled <- penguins |> na.omit() |>
  mutate(s_flipper_length = scale(flipper_length_mm), 
         s_bill_length = scale(bill_length_mm),
         s_bill_depth = scale(bill_depth_mm),
         s_body_mass = scale(body_mass_g))
```
```{r echo = FALSE, fig.align='center', out.width="70%", fig.asp = .6}
penguins_scaled |> ggplot() +
  geom_density(aes(s_flipper_length, color = "Flipper length")) + 
  geom_density(aes(s_bill_length, color = "Bill length")) + 
  geom_density(aes(s_bill_depth, color = "Bill depth")) + 
  geom_density(aes(s_body_mass, color = "Body mass")) + 
  theme(aspect.ratio = .6)
```

## Perform k-means

```{r}
set.seed(1)
kclust1 <- kmeans(penguins_scaled |>
                    select(s_flipper_length, s_body_mass, s_bill_length, s_bill_depth),
                  centers = 3)
tidy(kclust1) |> kable(digits = 3)
```

## Display clustering

```{r fig.align="center", out.width="60%", fig.asp=0.8}
kclust1 |> augment(penguins_scaled) |>
  ggplot(aes(y = s_bill_depth, x = s_bill_length, color = .cluster, shape = species)) + 
  geom_point() + theme(aspect.ratio = .8)
```

## Alternate display 

```{r fig.align="center", out.width="60%", fig.asp=0.8}
kclust1 |> augment(penguins_scaled) |>
  unite("spp_clust", species, .cluster) |>
  ggplot(aes(y = s_bill_depth, x = s_bill_length, color = spp_clust)) + 
  geom_point() + theme(aspect.ratio = .8)
```


## How good is the clustering?

```{r echo = FALSE, out.width="60%", fig.align="center"}
kclusts2 <- 
  tibble(k = 1:9) |>
  mutate(
    kclust = map(k, ~kmeans(penguins_scaled |> select(s_bill_depth, s_bill_length, s_body_mass, s_flipper_length),
                            centers = .x)),
    tidied = map(kclust, tidy),
    glanced = map(kclust, glance),
    augmented = map(kclust, augment, penguins_scaled)
  )
clusters2    <- kclusts2 |> unnest(cols = c(tidied))
assignments2 <- kclusts2 |> unnest(cols = c(augmented))
clusterings2 <- kclusts2 |> unnest(cols = c(glanced))
clusterings2 |> mutate(proportion_within = tot.withinss / totss) |>
  ggplot(aes(x = k, y = proportion_within)) + 
  geom_line() + geom_point() + theme_bw() +
  scale_x_continuous(breaks = 1:9) + ylim(0,1)
```

## Four species of penguins?

```{r echo = FALSE}
kclust1 <- kmeans(penguins_scaled |>
                    select(s_flipper_length, s_body_mass, s_bill_length, s_bill_depth),
                  centers = 4)
# tidy(kclust1) |> kable(digits = 3)
kclust1 |> augment(penguins_scaled) |>
  ggplot(aes(y = s_bill_depth, x = s_bill_length, color = .cluster, shape = species)) + 
  geom_point() + theme(aspect.ratio = .8)
```

## When to use k-means?

* When the goal is to make categories from continuous variables

* Each observation is placed in a cluster

* Based on distance between observations, calculated across all variables

  * Usually best to scale (and maybe transform in other ways) each variable

* Select the number of clusters based on partitioning of "sums of squares"


## Further reading

* Course notes

* [How does k-means clustering work?](https://saturncloud.io/blog/what-is-k-means-clustering-and-how-does-its-algorithm-work/)


## Task

* Practice k-means



