---
title: "Dynamic graphics"
author: "Andrew Irwin, a.irwin@dal.ca"
date: "2022-11-03"
format:
  revealjs:
    slide-number: true
    theme:  default
    chalkboard: true
    auto-animate: true
    scrollable: true
    code-block-border-left: true
    code-copy: false
    code-link: true
    history: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) # testing
library(tidyverse)
library(here)
library(janitor)
library(plotly)
library(gganimate)
library(palmerpenguins)
my_theme = theme_classic() + 
  theme(text = element_text(size=18))
```

## Plan

* Advantages and disadvantages of dynamic graphics?

* When should you use dynamic graphics?

* Examples

* Application in this course

## When and Why should you use dynamic graphics?

* Interactivity

* Show changes over time (in the data) over time (as perceived by the viewer)

* Easy to make too complicated

* Requires interaction and may not immediately make the point you want to make

* Distracting

## Highlights and interaction

[Link to interactive graphic](dynamic-graphics-1.html)

```{r warning=FALSE}
p1 <- penguins |> plot_ly(x = ~ body_mass_g, y = ~ bill_length_mm) |>
  add_markers(color = ~ species)
p1
htmlwidgets::saveWidget(as_widget(p1), "dynamic-graphics-1.html")
```

## Animations

Easy to create with `gganimate`. Make a regular ggplot, then use a variable to show how it changes over time.

```{r warning=FALSE, fig.align="center", out.height="50%", out.width="50%"}
penguins |> ggplot(aes(bill_length_mm, bill_depth_mm,
                        color = species)) + geom_point() + my_theme
```

## Animations

Easy to create with `gganimate`. Make a regular ggplot, then use a variable to show how it changes over time.

Shown on next slide.

```{r warning=FALSE}
b = seq(from = 2500, to = 6500, by = 500)
anim1 <- penguins |> 
  mutate(size_class = cut(body_mass_g, breaks=b, dig.lab=4),
         group=1:n()) |>
  ggplot(aes(bill_length_mm, bill_depth_mm,
             color = species, group=group)) + 
  geom_point() + 
  labs(title = "Body mass in the interval {closest_state}") +
  transition_states(size_class) +
  enter_fade() + exit_fade() + my_theme
```

## Animations

```{r echo=FALSE, warning=FALSE, fig.align="center", fig.asp=1}
animate(anim1)
```

## Summary

* Dynamic and interactive graphics can be fun to create

* Making good use of these features requires practice

* Use sparingly! Think of your audience and goals

* Good examples: 
  * Rosling's animation of gapminder data over years
  * [blue whale avoiding ships](https://www.theverge.com/2021/2/3/22265059/animation-blue-whales-ship-strikes)

## A version of Rosling's graph

```{r}
animation <- gapminder::gapminder |> 
  ggplot() +
  geom_text(aes(label = format(round(year))), 
            x = 3.8, y = 50, 
            size = 40, color = "lightgray") +
  geom_point(aes(x = gdpPercap,
             y = lifeExp,
             size = pop,
             color = continent)) +
  theme_bw() +
  scale_x_continuous(trans = "log2") +
  scale_size_continuous(trans = "log10") +
  transition_time(year) +
  labs(title = "Year {frame_time}", x = "GDP per capita ($, log scale)", y = "Life Expectancy", size = "Population", color = "Continent")
```

## A version of Rosling's graph

```{r echo=FALSE, warning=FALSE, fig.align="center", fig.asp=0.65}
animate(animation, nframes = 75)
```




