---
title: "Making outline maps"
author: "Andrew Irwin, a.irwin@dal.ca"
date: "2026-03-24"
format:
  revealjs:
    slide-number: true
    theme:  default
    chalkboard: true
    auto-animate: true
    scrollable: true
    code-block-border-left: true
    code-copy: false
    code-link: true
    history: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)
# library(plotly)
# library(gganimate)
# library(palmerpenguins)
my_theme = theme_classic() + 
  theme(text = element_text(size=18))
```

## Plan

* Kinds of maps

* Basic maps

* Adding points

* Shading areas

* Projections

## Map libraries

* `sf` for "simple (geographic) features`
* `maps`
* `rnaturalearth` and `rnaturalearthdata` (see naturalearthdata.com)
* `cancensus` for detailed maps of Canada
* `mapproj` for making projections of the Earth's surface
* `leaflet` for drawing raster or tiled maps (next lesson)

## Kinds of maps

* Coastlines and political boundaries

* Natural features (rivers, water bodies)

* Points and filled regions on maps

* Tiled maps

How to represent latitude and longitude from a sphere on a flat screen? (Projections.)

## A basic map

```{r message=FALSE, warning=FALSE}
library(tidyverse, quietly = TRUE)
library(sf)
library(maps)
WorldData <- maps::map(plot=FALSE, fill=TRUE)  |> st_as_sf()
m1 <- ggplot(WorldData) + 
  geom_sf(fill = "grey", color = "white", linewidth = 0.2) +
  theme_bw()
```


## A basic map

```{r echo=FALSE, fig.asp=0.5, out.width="90%", fig.align="center"}
m1
```


## Colour countries

```{r message=FALSE, warning=FALSE}
WorldData <- WorldData |> mutate(value = runif(n()))
m2 <- WorldData |> ggplot() + 
        geom_sf(aes(fill = value),
                colour = "white", 
                linewidth=0.2) +
   theme_minimal()
```

## Colour countries

```{r out.width="90%", fig.asp = 0.5, echo=FALSE, fig.align="center"}
m2
```


## Show only some countries

```{r message=FALSE, warning=FALSE}
m3 <- WorldData |> 
  filter(ID %in% c("Canada", "China", "Chile")) |>
  ggplot() + 
    geom_sf(aes(fill = value),
             colour = "lightgray", 
             linewidth=0.2) +
  theme_void()
```

## Show some countries

```{r out.width="90%", fig.asp = 0.5, echo=FALSE, fig.align="center"}
m3
```


## Colour some countries

```{r message=FALSE, warning=FALSE}
values <- tibble( ID = c("Canada", "China", "Chile"),
                  value2  = c(1, 2, 3))
m4 <- WorldData |>
  left_join(values) |>
  ggplot() + 
       geom_sf(colour = "white", fill = "tan",
               linewidth=0.1) +
       geom_sf(aes(fill = value2), color = "white",
               linewidth = 0.1) + 
  scale_fill_continuous(na.value = "transparent") +
  theme_minimal()
```

## Colour some countries

```{r fig.asp = 0.6, out.width = "90%", echo=FALSE, fig.align="center"}
m4
```

## Add points

```{r message=FALSE, warning=FALSE}
library(rnaturalearth)
library(rnaturalearthdata)
library(ggrepel)
canada <- ne_countries(country = "canada", scale = "large", 
                       returnclass = "sf")
sites <- tibble(longitude = c(-81, -80), 
                latitude = c(44, 45), 
                label = c("A", "B"))
m5 <- ggplot(data = canada) +
      geom_sf() +
      geom_point(data = sites, 
                 aes(x = longitude, y = latitude), 
                 size = 4, shape = 21, fill = "darkred") +
  geom_label_repel(data = sites, 
                   aes(x = longitude, y = latitude, 
                       label = label)) + 
  coord_sf(xlim = c(-140,-50), ylim = c(42,65))
```

## Add points

```{r fig.asp = 0.6, out.width = "90%", echo=FALSE, fig.align="center"}
m5
```

## Projections

```{r message=FALSE, warning=FALSE}
m6 <- canada |> ggplot() + 
        geom_sf(fill = "lightgray", colour = "#7f7f7f") +
  theme_bw() +
  coord_sf(crs = st_crs(3978))  # Albers Conic Equal Area
```

## Projections

```{r fig.asp = 0.6, out.width = "90%", echo=FALSE, fig.align="center"}
m6
```

## Polar regions

```{r message=FALSE, warning=FALSE}
m7 <- WorldData |>
  ggplot() + 
        geom_sf(fill = "lightgray", colour = "#7f7f7f", 
                 alpha = 0.5, linewidth=0.2) +
  theme_bw() +
  coord_sf(crs = paste0("+proj=nsper +h=", 2.5 * 6371000, " +lat_0=60 +lon_0=-100"))
```

## Polar regions

```{r fig.asp = 0.6, out.width = "90%", echo=FALSE, fig.align="center"}
m7
```

## Other pole

```{r message=FALSE, warning=FALSE}
m7a <- WorldData |>
  ggplot() + 
        geom_sf(fill = "lightgray", colour = "#7f7f7f", 
                 alpha = 0.5, linewidth=0.2) +
  theme_void() +
  coord_sf(crs = paste0("+proj=nsper +h=", 2.5 * 6371000, " +lat_0=-75 +lon_0=0"))
```

## Polar regions

```{r fig.asp = 0.6, out.width = "90%", echo=FALSE, fig.align="center", warning=FALSE}
m7a
```

## Map of Canada

With provinces and territories

```{r}
canada <- ne_states(country = "canada", returnclass = "sf")
m8 <- canada |> ggplot() + 
        geom_sf(fill = "lightgray", colour = "#7f7f7f") +
    coord_sf(crs = st_crs(3347)) +  # Lambert Conformal
theme_void() +
  theme(
    panel.background = element_rect(fill = "aliceblue"),
    panel.grid = element_blank()
  )
```

---

```{r fig.asp = 0.6, out.width = "90%", echo=FALSE, fig.align="center", warning=FALSE}
m8
```

## Map of Canada with Census divisions

```{r}
library(cancensus)
# set_cancensus_cache_path('/tmp', install = TRUE)
# https://censusmapper.ca/
# https://mountainmath.github.io/cancensus/index.html#api-key
canada_counties <- get_census(
  dataset = "CA21", 
  regions = list(C = "01"), 
  level = "CD", 
  geo_format = "sf"
)
m9 <- ggplot(data = canada_counties) +
  geom_sf(fill = "white", color = "gray50", size = 0.1) +
  coord_sf(crs = 3347) + 
  theme_minimal()
```

---

```{r fig.asp = 0.6, out.width = "90%", echo=FALSE, fig.align="center", warning=FALSE}
m9
```

## Population density

```{r}
# v_CA21_6 is the vector for Population Density in the 2021 Census
# You can find vector IDs using: list_census_vectors("CA21")
pop_data <- get_census(
  dataset = "CA21", level = "CD",
  regions = list(C = "01"),
  vectors = c("pop_density" = "v_CA21_6"),
  geo_format = "sf")
m10 <- ggplot(pop_data) +
  geom_sf(aes(fill = pop_density), color = NA) +
  scale_fill_viridis_c(
    name = "Pop Density", 
    trans = "log10", option = "magma") +
  coord_sf(crs = 3347) +
  theme_void() +
  labs(title = "Population Density by Census Division")
```


---

```{r fig.asp = 0.6, out.width = "90%", echo=FALSE, fig.align="center", warning=FALSE}
m10
```


## Summary

* Make a basic map

* Select all or some countries

* Fill regions with colour

* Add points and labels

* Use a different projection


## Exercises

* Make some simple maps following the examples in these slides and the course notes




