---
title: "Making outline maps"
author: "Andrew Irwin, a.irwin@dal.ca"
date: "2024-03-19"
format:
  revealjs:
    slide-number: true
    theme:  default
    chalkboard: true
    auto-animate: true
    scrollable: true
    code-block-border-left: true
    code-copy: false
    code-link: true
    history: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)
library(plotly)
library(gganimate)
library(palmerpenguins)
my_theme = theme_classic() + 
  theme(text = element_text(size=18))
```

## Plan

* Kinds of maps

* Basic maps

* Adding points

* Shading areas

* Projections

## Map libraries

* `ggmap` (part of tidyverse)

* Many other libraries 
  * `maps`
  * `mapproj` for making projections of the Earth's surface
  * `mapdata` with maps of China and Japan
  * `leaflet` for drawing raster or tiled maps (next lesson)

## Kinds of maps

* Coastlines and political boundaries

* Natural features (rivers, water bodies)

* Points and filled regions on maps

* How to represent latitude and longitude from a sphere on a screen?

* Tiled maps

## A basic map

```{r message=FALSE, warning=FALSE}
library(tidyverse)
WorldData <- map_data('world')  
m1 <- ggplot(WorldData) + 
        geom_map(map = WorldData,
                 aes(map_id = region),
                 fill = "lightgray", colour = "#7f7f7f", 
                 alpha = 0.5, linewidth=0.5) +
  expand_limits(x = c(-180, 180), y = c(-85, 85)) + 
  theme_bw()
```


## A basic map

```{r echo=FALSE, fig.asp=0.5, out.width="90%", fig.align="center"}
m1
```


## Colour countries

```{r message=FALSE, warning=FALSE}
values <- tibble( region = WorldData |> pull(region) |> unique())
values <- values |> mutate(value = runif(nrow(values)))
m2 <- values |> ggplot() + 
        geom_map(aes(fill = value, map_id = region),
                 map = WorldData,
                 colour = "lightgray", 
                 alpha = 1, linewidth=0.5) +
   expand_limits(x = c(-180,180), y = c(-80,80)) +
   theme_bw()
```

## Colour countries

```{r out.width="90%", fig.asp = 0.5, echo=FALSE, fig.align="center"}
m2
```


## Show only some countries

```{r message=FALSE, warning=FALSE}
values <- tibble( region = c("Canada", "China", "Chile"),
                  value  = c(1, 2, 3))
m3 <- values |> ggplot() + 
        geom_map(aes(fill = value, map_id = region),
                 map = WorldData,
                 colour = "lightgray", 
                 alpha = 1, linewidth=0.5) +
   expand_limits(x = c(-180,180), y = c(-80,80))+
  theme_bw()
```

## Show some countries

```{r out.width="90%", fig.asp = 0.5, echo=FALSE, fig.align="center"}
m3
```


## Colour some countries

```{r message=FALSE, warning=FALSE}
values <- tibble( region = c("Canada", "China", "Chile"),
                  value  = c(1, 2, 3))
m4 <- ggplot(data = WorldData, aes(map_id = region)) + 
        geom_map(map = WorldData,
                 colour = "lightgray", fill = "darkgray",
                 alpha = 1, linewidth=0.5) +
        geom_map(aes(fill = value), 
                 data = values, map = WorldData) + 
   expand_limits(x = c(-180,180), y = c(-80,80))+
  theme_bw()
```

## Colour some countries

```{r fig.asp = 0.6, out.width = "90%", echo=FALSE, fig.align="center"}
m4
```

## Add points

```{r message=FALSE, warning=FALSE}
library(rnaturalearth)
library(rnaturalearthdata)
library(ggrepel)
canada <- ne_countries(country = "canada", scale = "large", 
                       returnclass = "sf")
sites <- tibble(longitude = c(-81, -80), latitude = c(44, 45), 
                label = c("A", "B"))
m5 <- ggplot(data = canada) +
      geom_sf() +
      geom_point(data = sites, 
                 aes(x = longitude, y = latitude), 
                 size = 4, shape = 21, fill = "darkred") +
  geom_label_repel(data = sites, 
                   aes(x = longitude, y = latitude, 
                       label = label)) + 
  coord_sf(xlim = c(-140,-50), ylim = c(42,65))
```

## Add points

```{r fig.asp = 0.6, out.width = "90%", echo=FALSE, fig.align="center"}
m5
```

## Projections

```{r message=FALSE, warning=FALSE}
library(mapproj)
CanadaData <- map_data('world', 'Canada')  
m6 <- ggplot(CanadaData, aes(map_id=region)) + 
        geom_map(map = CanadaData,
                 aes(x = long, y = lat),
                 fill = "lightgray", colour = "#7f7f7f", 
                 alpha = 0.5, linewidth=0.5) +
  theme_bw() +
  coord_map("lambert", lat0 = 42, lat1 = 50)
```

## Projections

```{r fig.asp = 0.6, out.width = "90%", echo=FALSE, fig.align="center"}
m6
```

## Polar regions

```{r message=FALSE, warning=FALSE}
library(mapproj)
m7 <- ggplot(WorldData, aes(map_id=region)) + 
        geom_map(map = WorldData,
                 aes(x = long, y = lat),
                 fill = "lightgray", colour = "#7f7f7f", 
                 alpha = 0.5, linewidth=0.5) +
  theme_bw() +
  coord_map("perspective", 2, orientation=c(-90, -90, 90))
```

## Polar regions

```{r fig.asp = 0.6, out.width = "90%", echo=FALSE, fig.align="center"}
m7
```

## Summary

* Make a basic map

* Select all or some countries

* Fill regions with colour

* Add points and labels

* Use a different projection


## Task

* Bonus task - make some simple maps based on course notes




