---
title: "Graphics output"
author: "Andrew Irwin, a.irwin@dal.ca"
date: "2024-04-02"
format:
  revealjs:
    slide-number: true
    theme:  default
    chalkboard: true
    auto-animate: true
    scrollable: true
    code-block-border-left: true
    code-copy: false
    code-link: true
    history: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(forcats)
library(lubridate)
library(palmerpenguins)
library(kableExtra)
library(paletteer)
library(gapminder)
my_theme = theme_bw() + theme(text = element_text(size=25))
knitr::opts_chunk$set(
  echo = FALSE,
  fig.width = 6, fig.asp = 0.65, fig.align="center", out.width = '70%', dev='svglite'
)
```

## Plan

* Kinds of graphics 

* How to describe the size of a graphic

* Knitted documents and separate files


## Starting figure

```{r}
p1 <- gapminder |> # filter(continent == "Europe") |>
  ggplot(aes(x = year, y = lifeExp)) + 
  # geom_line(aes(group=country)) + 
  geom_smooth(method = "loess", formula = y ~ x, aes(color = continent)) +
  theme_bw() + 
  theme(text = element_text(size=20))
p1
```

## Kinds of graphics

* Vector graphics (pdf, svg)

* Bitmapped or raster graphics (png, jpeg)

* Why does it matter?
  
  * File sizes
  
  * Quality when printed
  
  * Appearance on screen
  
## Overview

* Device types: png, ragg, jpeg, svg, pdf

* Size: fig.height, fig.width (in inches, default 7), and fig.asp (aspect ratio: height / width)

* Display size: out.width, out.height (in percentage of display area, e.g., '70%')

* Resolution:  dpi (dots per inch; 96-250 on screen, 300-600 when printed)

* Alignment: fig.align = left, right, center

## Chunk options

R markdown:

````
```{{r fig.height = 5, fig.asp = 0.6}}
# code for plot here
```
````

Quarto document (as above, or...)

```{{r}}
#| fig.height: 5
#| fig.asp: 0.5
# code for plot here
```
   

## PNG example

Chunk options: `fig.width = 6, fig.asp = 0.65, fig.align="center", out.width = '70%', dev = 'png'`

```{r dev='png'}
p1 
```

## Higher resolution PNG example

Chunk options: `dev='png', dpi=150`

```{r dev='png', dpi=150}
p1 
```

## RAGG example

Chunk options: `dev='ragg_png', dpi = 150`

```{r dev='ragg_png', dpi = 150}
p1 +
  annotate("text", 1990, 45, label = "Tokyo (東京)") + 
  labs(title = "Text in Japanese (日本) 🇯🇵") 
```


## Fails with PNG

Chunk options: `dev='png', dpi = 150`

```{r dev='png', dpi = 150}
p1 +
  annotate("text", 1990, 60, label = "Tokyo (東京)") + 
  labs(title = "Text in Japanese (日本) 🇯🇵") 
```

## JPEG example

Chunk options: `dev='jpeg', dpi = 150`

```{r dev='jpeg', dpi = 150}
p1 
```

## SVG example

Chunk options: `dev='svglite'`

```{r dev='svglite'}
p1 
```

## Creating standalone files

Use `ggsave`:

```{r eval=FALSE, echo=TRUE}
ggsave("test-plot.pdf", p1, height = 4, width = 6)
ggsave("test-plot.png", p1, dpi = 400)
```





